---
title: "A Bayesian model using Hamiltonian Monte Carlo sampling built in R and evaluated by TensorFlow"
subtitle: "Using greta::mcmc()" 
output: html_notebook
---

```{r}

install_tensorflow()
install.packages(c("coda","mvtnorm","devtools","loo"))
#library(devtools)
#devtools::install_github("rmcelreath/rethinking")
library(rethinking)
library(greta)
library(DiagrammeR)
library(bayesplot)
library(dplyr)
library(ggplot2)

```


Prepare data
```{r}
# Example from section 8.3 Statistical Rethinking

data(rugged)
# d <- rugged
# d$log_gdp <- log(d$rgdppc_2000)
# dd <- d[complete.cases(d$rgdppc_2000), ]
# dd_trim <- dd[ , c("log_gdp","rugged","cont_africa")]

dd_trim <- rugged %>% # d %>% 
  mutate(log_gdp = log(d$rgdppc_2000)) %>% 
  filter(complete.cases(rgdppc_2000)) %>% 
  select(log_gdp, rugged, cont_africa) 

dd_trim

#set.seed(1234)

```


# Set up tensor flow's data structures (greta's data_node array objects)
Note: greta doesn’t immediately calculate values for new arrays. It works out the size and shape of the result and creates a place-holder data structure.

```{r}

g_log_gdp <- as_data(dd_trim$log_gdp)
g_rugged <- as_data(dd_trim$rugged)
g_cont_africa <- as_data(dd_trim$cont_africa)

```


# Set up Bayesian model
Note: The parameters a, bR, bA, bAR, sigma, and mu are all new greta arrays that don’t contain any data
```{r}

# Variables and Priors

a <- normal(0, 100)
bR <- normal(0, 10)
bA <- normal(0, 10)
bAR <- normal(0,10)
sigma <- cauchy(0,2,truncation=c(0,Inf))

# > a
# 
# greta array (variable following a normal distribution)
# 
#      [,1]
# [1,]  ?  

# operations
mu <- a + bR*g_rugged + bA*g_cont_africa + bAR*g_rugged*g_cont_africa

# > dim(mu)
## [1] 170   1

# sets up the likelihood function for the model
distribution(g_log_gdp) = normal(mu, sigma)

```


Define the model 
Note: The model() function does all of the work. It fits the model and produces a fairly complicated object organized as three lists that contain, respectively, the R6 class, TensorFlow structures, and the various greta data arrays.
```{r}

# defining the model
mod <- model(a,bR,bA,bAR,sigma)

str(mod,give.attr=FALSE,max.level=1)

# plotting
plot(mod)
```


Use greta::mcmc() to sample from the posterior distributions defined in the model

```{r}
draws <- mcmc(mod, n_samples = 1000)
summary(draws)
```


Posterior distribution of the interaction term (bAR)

```{r}

mat <- data.frame(matrix(draws[[1]],ncol=5))
names(mat) <- c("a","bR","bA","bAR","sigma")

mat %>% 
  ggplot(aes(x=bAR)) + 
  geom_histogram(aes(y=..density..), binwidth=.05, colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```




* mcmc_trace() to examine the trace plots for the MCMC samples using the greta function. 
  - The plots for each parameter appear to be stationary (flat, i.e., centered on a constant value) and well-mixed (there is no obvious correlation between points). 
* mcmc_intervals() plots the uncertainty intervals for each parameter computed from posterior draws with all chains merged.
```{r}
mcmc_trace(draws)

mcmc_intervals(draws)
```


