---
title: "Multithreading and Map-Reduce in Stan 2.18.0: A Minimal Example"
version: "v1 (19 Sep 2018)"
output: html_notebook
---

# Multithreading and Map-Reduce in Stan 2.18.0: A Minimal Example
## v1 (19 Sep 2018)

This is a modified version of 
the beginner tutorial for using cmdstan with multithreading by Richard McElreath: https://github.com/rmcelreath/cmdstan_map_rect_tutorial 

# Load packages
```{r}
library(here)
library(magrittr)
library(tidyverse)

```

# List converted from data read from github
```{r}
d <- RCurl::getURL("https://raw.githubusercontent.com/rmcelreath/cmdstan_map_rect_tutorial/master/RedcardData.csv") %>% 
  read.csv(text = ., stringsAsFactors=FALSE )
table( d$redCards )  # build a contingency table of the counts at each combination of factor levels

out_data <- d %>% 
  filter(!is.na(rater1)) %>% 
  sample_n(9999) %>% 
  select(n_redcards = redCards, n_games = games, rating = rater1) 
  
out_data <- list(N = out_data %>% nrow(), as.list(out_data)) %>% flatten()

library(rstan)
options(mc.cores = parallel::detectCores()) # For execution on a local, multicore CPU with excess RAM 
rstan_options(auto_write = T)  # avoid recompilation of unchanged Stan programs
#Sys.setenv(LOCAL_CPPFLAGS = '-march=native') # For improved execution time

```


# Baseline example of Stan model
```{r}

writeLines(readLines("McElreath_eg.stan"))

```

# Fit the baseline model from the .stan file
```{r}

m0 <- stan(file = 'McElreath_eg.stan', data=out_data, seed=7654321, refresh=1000, chains = 1)  


```

# Stan code chunk specifying the model (with a stanfit object as the output)
```{stan output.var = McElreath_eg}

// .stan for McElreath_Example_Multithreading.Rmd

 data {
  int N;
  int n_redcards[N];
  int n_games[N];
  real rating[N];
}

parameters {
  vector[2] beta;
}

model{
  beta ~ normal(0,1);
  n_redcards ~ binomial_logit( n_games, beta[1] + beta[2] * to_vector(rating) );
}

```

# Fit the baseline model from the stanfit object created from the Stan code chunk
```{r}

m1 <- stan(model_code = McElreath_eg@model_code, data=out_data, seed=7654321, refresh=1000, chains = 1)  


```
