---
title: "State space model of accounting manipulation - Stan"
output: html_notebook
fig_width: 6 
fig_height: 6 
---

```{r}

#knitr::purl("ssm_Stan0.5.2.Rmd", output = "ssm_Stan0.5.2.R", documentation = 2)
.libPaths()
.libPaths( c( "~/R_library/3.5", .libPaths() ) )
#.libPaths( c( .libPaths(), "~/R_library/3.5" ) )
.libPaths()

#library(tidyverse)
library(dplyr)
library(magrittr)

#load previously saved Stan model objects
#load(file = "StanObjects.rda")

#library(shinystan)
library(bayesplot)
library(rstan)
rstan_options(auto_write = TRUE) # avoid recompilation of unchanged Stan programs

# The following can affect distributional computing over a cluster
options(mc.cores = parallel::detectCores())  # 

# The following throws an error in compiling .stan
#Sys.setenv(LOCAL_CPPFLAGS = '-march=native') # For improved execution time but can throw errors for some CPUs

```


# Estimate the simulated model with Stan - misreporting extent 
  (forecasts specified in generated quantities block)
//ignore these lines: 
//  COGS2Rev, SGA2Rev, RnD2Rev with Rev2Rev = unit vector as the intercept)
//   int<lower=0,upper=1> M[N];   // M = 1 for the decision to misreport; = 0 if report honestly  

```{stan output.var=SSM, eval=F}
```

# Simulate data
```{stan output.var=simu_SSM, eval=F}
```



# Simulate data and fit the debug model (also, save Stan model objects) 
```{r}
# J=80,N=4*30: Apparently, no need to go beyond this (take WRDS about 2.5 hrs to finish)
# J=45,N=4*30 gives reasonable estimates of the baseline model of AR(1) with hierarchical seasonability
# J=60,N=4*20 also seems to be acceptable
# Can consider J=42,N=4*18 for model dev process only  
#   but J=45,N=4*20 gives more reasonable estimates  
#===========================================
J = 90#60#42#40#45#20 #200     # number of firms
N = 4*25#30#18#15#5#15   #number of (quarterly) observations of a firm
Q = 4 # quarterly time series

rho = 1#2#5  # curvature of the soft-clipping function; rho=50 would make nearly hard-clipping
mu_base = -1.6
sd_base = 0.2 #0.05  
mu_alpha = 0.04 #0.1#0.2  # intercept parameter of the AR(1) processs of the underlying unbiased figure
#mu_alpha = 0.04  # intercept parameter of the AR(1) processs of the underlying unbiased figure
beta = 0.6#0.85  # slope parameter of the AR(1) processs of the underlying unbiased figure
theta = 0.15 #0.2 #0.1 # real EM's LT impact on alpha[j,n]
y_LT = mu_alpha/(1-beta)
mu_u1 = y_LT # initial y for data simulation set to three times of the LT stationary level
ab_mu = c(mu_alpha, beta)
sd_y = 0.08  # sd of underlying unbiased figures is likely to be industry-specific / business model specific
mu_season = c(-0.12, -0.06, 0.15)
sd_season = 0.1

  
  #delta = 0.65 # reversal coeff. of real EM (eg, prior year's heavily discounted sales taking from this year's)
  # delta can be bigger than, eg, sales taken from last year's heavy discount that would have been higher w/o discount; but delta can be less than 1 if the heavy discount does stimulate more sales than taken from next yaer's

#pvec <- matrix(data = p, nrow = I)  # define column vector as a matrix[N,1]
#Z.l[[3]]%*%p %>%  str()  #ok with p as will be converted accordingly


# parameters of the DGP of the likelihood of using sales- vs. RnD-based real EM (sigma = 1 or 0)
s = c(0.4, 0.15, 0.15) # s = c(0.4, 0.15, 0.15) 
S = length(s)

# parameters of the data generation process of the temptation to do real EM
#p = c(0.72, 0.42)#, 0.7) # p[3] is constant and hard-coded; this p[3] sets Real|p[1:2]=0 at a reasonable level 
p = c(0.7, 0.62, 0.68) # for inv_logit(): changed p[3] to see if it could be reliably est'd
#p = c(0.7, 0.62, 0.35) # p[1] is constant; these parameters are for inv_logit()
# p = c(0.7, 0.72, 0.42) # p[1] is constant; these parameters are for soft-clipping function
# each factor contributes only 1-1.1% but both together can go up to 16% pt. of real EM
I = length(p) # soft-clipping  rho=5

# parameters of the data generation process of the temptation to misreport
g = c(0.85, 0.35) 
# g = c(0.2, 0.2, 0.25) 
K = length(g)

# parameters of the data generation process of the strength of governance and monitoring 
# (eg, audit committee chair with financial expertise, big 4 auditor, etc)
w = c(0.72, 0.15, 0.62) # w = c(0.72, 0.15, 0.62) 
H = length(w)

d = c(0.05, 0.7, 0.25) # fractional reversal of prior-period manipluation by accruals
L = length(d)

NoReportedRnD_freq = 0.85
IsRetailer_freq = 0.55
ratioNED_req = 0.25 #0.3  # min ratio of non-executive directors (NED) required by law
Big4_freq = 0.7#0.6
downgrade_freq = 0.3 #0.5
integrity = 0.4 #0.35 #(0,1)  0.5 = average       
# integrity level, between (0,1), of the society, interpreted as likelihood to behave opportunistically
# integrity level of an individual, indiv_integ between (-Inf,Inf), is 
#   affected by a society's general integrity level in a stochastic fashion varying from period to period

#M <- rbinom(N, 1, integrity)   # the decision to misreport earnings

# Define function to convert data.frame with gvkey as the key to a list of matices
df2mat.l <- function(z){
  z %>% 
  split(.$gvkey) %>% 
  lapply(function(z) select(z, -c(gvkey))) %>% 
  lapply(as.matrix)
  }

#set.seed(8)
# T.l captures firm characteristics related to whether firms are more likely to use sales- vs. RnD-based real EM
T.l <- data.frame(
  gvkey = rep(1:J, times=N),                   
  constant = rep(1, J*N),
  NoReportedRnD = (10)*( rbinom(J*N, 1, NoReportedRnD_freq) - NoReportedRnD_freq ),
  IsRetailer = (10)*( rbinom(J*N, 1, IsRetailer_freq) - IsRetailer_freq )#,
 ) %>%
  df2mat.l()

# make Z.l different from X.l to avoid less than full rank in estimation
Z.l <- data.frame(
  gvkey = rep(1:J, times=N),                   
  constant = (-10)*rep(1, J*N),
  distortion = (10)*( rbeta(J*N, 3*(1-integrity), 3*integrity) - (1-integrity) ),
  downgrade = (10)*( rbinom(J*N, 1, downgrade_freq) - downgrade_freq )#,
#constant = (-10)*rep(1, J*N)#,
 ) %>%
  df2mat.l()

# covariate matrix X capturing the dis/incentive to misreport
# downgrade = 1 if one or more analysts following the firm downgraded their stock recommendation in the last period
X.l <- data.frame(
  gvkey = rep(1:J, times=N),                   
  constant = rep(1, J*N),
  aggressive = (10)*( rbeta(J*N, 3*(1-integrity), 3*integrity) - 0.4 )#,  #(1-integrity)
#  downgrade = (10)*rbinom(J*N, 1, downgrade_freq)#,
  ) %>%
  df2mat.l()

get_ratioNED <- function(ratioNED_req, J, N) {
  CC = pmax(ratioNED_req, rbinom(J, 20, ratioNED_req)/20) 
  EE = rep(NULL, J)
  for (n in 1:N)  { 
    A = rbinom(J, 1, 0.5)/20
    B = -rbinom(J, 1, 0.5)/20
    CC = pmin(0.65, pmax(ratioNED_req, A+B+CC)) #%>%  print()
    EE = c(EE, CC)
    } 
  return(EE)
  }
# covariate matrix G capturing the strength of internal governance and external monitoring mechansims
G.l <- data.frame(
  gvkey = rep(1:J, times=N), # note the difference from times=N
  constant = (1)*rep(1, J*N),
  ratioNED = (10)*get_ratioNED(ratioNED_req, J, N), 
  Big4 = (2)*rbinom(J*N, 1, Big4_freq)#,
  ) %>%
  df2mat.l()
# G.l %>%  arrange(gvkey) %>% head(120)
#colnames(Z) <- c("Rev2Rev", "COGS2Rev", "SGA2Rev", "RnD2Rev")

simu_SSM_data <- list(J = J, 
                      N = N, 
                      Q = Q,
                      S = S,
                      I = I,
                      K = K,
                      H = H,
                      L = L,
                      T = T.l,
                      Z = Z.l,
                      X = X.l,
                      G = G.l,
rho = rho,
                      sd_y = sd_y,
                      mu_season = mu_season,
                      sd_season = sd_season,
                      mu_u1 = mu_u1,
                      theta = theta,
                      mu_alpha = mu_alpha,
                      beta = beta, 
#sd_alpha = sd_alpha,
#                      sd_gamma = sd_gamma,
#                      sd_omega = sd_omega,
                      mu_base = mu_base,
                      sd_base = sd_base,
                      s = s,
                      p = p,
                      g = g,
                      w = w,
                      d = d
                      )
# save(simu_SSM_data, file="simu_SSM_data.rda")
# load(file="simu_SSM_data.rda")

# Simulate SSM data
simu_fit <- stan(#model_code = simu_SSM@model_code,
                 file = "simu_SSM0.5.2.stan",  # Stan program
#                 file = "simu_SSM2.4.stan",  # Stan program
                 data=simu_SSM_data, iter=1, #seed=987,  
                 chains=1, algorithm="Fixed_param")  

q <- extract(simu_fit)$q[1,] 

r.mat <- extract(simu_fit)$r[1,,]
y.mat <- extract(simu_fit)$y[1,,]
Real.mat <- extract(simu_fit)$Real[1,,]
m.mat <- extract(simu_fit)$m[1,,]
D.mat <- extract(simu_fit)$D[1,,]
b.mat <- extract(simu_fit)$b[1,,]
R.mat <- extract(simu_fit)$R[1,,]
sigma.mat <- extract(simu_fit)$sigma[1,,]

Real.mat[1,] %>% round(4)
Real.mat %>% round(4) %>%  mean()
Real.mat %>% round(4) %>%  median()
Real.mat %>% round(4) %>%  max()
Real.mat %>% round(4) %>%  min()
cat("\n\n")
r.mat[1,] %>% round(4)
y.mat[1,] %>% round(4)
y.mat[,1] %>% round(4) %>%  mean()
y.mat[,1] %>% round(4) %>%  max()
y.mat[,1] %>% round(4) %>%  min()

b.mat[1,] %>% round(4)
b.mat %>% round(4) %>%  max()
b.mat %>% round(4) %>%  min()

R.mat[1,] %>% round(4)
R.mat %>% round(4) %>%  max()
R.mat %>% round(4) %>%  min()
cat("\n\n")
m.mat[1,] %>% round(4)
D.mat[1,] %>% round(4)
D.mat %>% round(4) %>%  mean()
D.mat %>% round(4) %>%  max()
D.mat %>% round(4) %>%  min()
cat("\n\n")
sigma.mat[1,] %>% round(4)
sigma.mat %>% round(4) %>%  mean()
sigma.mat %>% round(4) %>%  median()
sigma.mat %>% round(4) %>%  max()
sigma.mat %>% round(4) %>%  min()

# Prepare SSM data for estimation using MCMC
#N_new = 3
SSM_data <- list(J = J,
                 N = N, 
                 Q = Q,
                 q = q,   
                 S = S,
                 I = I,
                 K = K,
                 H = H,
                 L = L,
#mu_base = mu_base,
rho = rho,
                 r = r.mat, 
                 T = T.l,
                 Z = Z.l,
                 X = X.l,
                 G = G.l#,
                 # Z_new = Z[1:N_new,],
                 # N_new = N_new
                 )

#stan_rdump(names(SSM_data), file = "SSM_data_J90c.dat", envir = list2env(SSM_data))

```

# Fit the full model (by referring to the debug model) to estimate underlying parameters
```{r}

#pdf("Rplots_FREEp1,1.5K,120,0.95,4ch.pdf")
n_iter = 2000#2000#3000
n_refresh = max(2, n_iter/10)
n_warmup = 1000
#n_warmup = n_iter/2 
n_chains = 4



initf2 <- function(chain_id = 1) {
  list(mu_alpha = mu_alpha, beta = beta, theta = theta, ab_mu = ab_mu, #shape = shape,
       s = s, p = p, g = g, w = w, d = d, mu_season = mu_season, sd_season = sd_season,
       mu_u1 = mu_u1, #u = array(y_LT, dim = c(J,N)),
       rho = rho, sd_y = sd_y, mu_base = mu_base, sd_base = sd_base) #, sd_gamma = sd_gamma, sd_omega = sd_omega)
  }
# generate a list of lists to specify initial values
init_ll <- lapply(1:n_chains, function(id) initf2(chain_id = id))

# Run the full model and refer to the debug model to save compilation time 
system.time({
fit2 <- stan(
#fit1 <- stan(
  
file = "SSM0.5.2.stan",  # Stan program
#  file = "SSM2.4.stan",  # Stan program
  data = SSM_data,    # named list of data
#  model_code = SSM@model_code, 
#  fit = fit0_debug,   # to save compilation time if the debug model was run
  control = list(adapt_delta = 0.8,#9,#8, #9
                 metric = "diag_e",#"dense_e",
#                 stepsize = 0.05, #0.05, #0.01
                 max_treedepth = 15),#15),    # adjust when there're divergent transitions after warmup
  init = init_ll, 
  chains = n_chains,#1,#8,             # default is 4 Markov chains
#  cores = 8,
#  seed = 555, 
  iter = n_iter,#4000,#7000,#12000, #500 #2000,            # total number of iterations per chain
  warmup = n_warmup,#500,          # number of warmup iterations per chain
  refresh = n_refresh#500          # = 0 means no progress shown
  )
})

#fit2 <- fit1
#save(fit2 , file="fit2.rda")#, file="/scratch/city/fit2.rda")
# beepr::beep()

#extract(fit1)$y[1,]
#pairs(extract(fit1), pars = c(a, b)) #, y[1], sd_y, m[1], sd_m, c[1], c[2], c[3]))

```

# Summary of the posterior sample 
```{r} 
#====================================================
  # sd_y = 0.08, mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1, # sd_base = 0.2; mu_base = -1.6 
  # sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0.4, 0.15, 0.15)  
  # p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.7, 0.72, 0.42) [soft-clipping]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.85, 0.35)       #old: g = c(0.2, 0.2, 0.25) 
  # w = c(0.72, 0.15, 0.62)  
# d = c(0.05, 0.7, 0.25) 
#====================================================

print(fit2, c(
              "sd_y", "mu_u1", "mu_alpha", "beta", #"rho", 
              "theta", "sd_season", "mu_season", #"sd_base", #"mu_base", 
#              "s",
              "p",
              "g", 
              "w", 
              "d", 
              "gw_mu", "gw_sd", "gw_L", "gw_err", 
              "p_mu", "p_sd", "p_L", "p_err", 
              "ab_mu", "ab_sd", "ab_L", "ab_err"
              ), 
      probs = c(0.1, 0.5, 0.9), digits = 3) #, probs = c(.05,.95))

get_num_leapfrog_per_iteration(fit2)

# fit <- read_stan_csv("samples.csv")
# print(fit, c(
#               "sd_y", "mu_u1", "mu_alpha", "beta", #"rho", 
#               "theta", "sd_season", "mu_season", #"sd_base", #"mu_base", 
# #              "s",
#               "p",
#               "g", 
#               "w", 
#               "d", 
#               "gw_mu", "gw_sd", "gw_L", "gw_err", 
#               "p_mu", "p_sd", "p_L", "p_err", 
#               "ab_mu", "ab_sd", "ab_L", "ab_err"
#               ), 
#       probs = c(0.1, 0.5, 0.9), digits = 3) #, probs = c(.05,.95))
# get_num_leapfrog_per_iteration(fit)

#====================================================
  # sd_y = 0.08, mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1, # sd_base = 0.2; mu_base = -1.6 
  # sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0.4, 0.15, 0.15)  
  # p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.7, 0.72, 0.42) [soft-clipping]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.85, 0.35)       #old: g = c(0.2, 0.2, 0.25) 
  # w = c(0.72, 0.15, 0.62)  
# d = c(0.05, 0.7, 0.25) 
#====================================================

```


# Pairs plots
```{r} 
posterior <- as.array(fit2)
lp <- log_posterior(fit2)
np <- nuts_params(fit2)  # nuts parameters
#par(mar = c(4, 4, 0.5, 0.5))
knitr::opts_chunk$set(fig.width=6, fig.height=6)
color_scheme_set("gray") #"darkgray")
#Error: cannot allocate vector of size 782.8 Mb
#mcmc_parcoord(posterior, np = np)

#http://mc-stan.org/bayesplot/articles/visual-mcmc-diagnostics.html
#The main problem is that large steps are required to explore the less narrow regions efficiently, but those steps become too large for navigating the narrow region. The required step size is connected to the value of tau. When tau is large it allows for large variation in theta (and requires large steps) while small tau requires small steps in theta.
#The non-centered parameterization avoids this by sampling the eta parameter which, unlike theta, is a priori independent of tau. Then theta is computed deterministically from the parameters eta, mu and tau afterwards. Here’s the same plot as above, but with eta[1] from non-centered parameterization instead of theta[1] from the centered parameterization: 


mcmc_pairs(posterior, np = np, pars = c("theta", "mu_season[1]", "mu_season[2]", "mu_season[3]"))

mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "theta", "mu_season[1]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "mu_season[2]", "mu_season[3]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "theta", "mu_season[1]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "mu_season[2]", "mu_season[3]"))

mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "theta", "mu_season[1]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "mu_season[2]", "mu_season[3]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[3]", "gw_mu[4]", "gw_mu[5]", "theta", "mu_season[1]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[3]", "gw_mu[4]", "gw_mu[5]", "mu_season[2]", "mu_season[3]"))

mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p_mu[3]", "theta", "mu_season[1]"))
mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p_mu[3]", "mu_season[2]", "mu_season[3]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "theta", "mu_season[1]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "mu_season[2]", "mu_season[3]"))

mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "ab_mu[1]", "ab_mu[2]")) 
mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "gw_mu[3]", "gw_mu[4]", "gw_mu[5]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "p_mu[1]", "p_mu[2]", "p_mu[3]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "d[1]", "d[2]", "d[3]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "p_mu[1]", "p_mu[2]", "p_mu[3]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "d[1]", "d[2]", "d[3]"))

#====================================================
  # sd_y = 0.08, mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1, # sd_base = 0.2; mu_base = -1.6 
  # sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0.4, 0.15, 0.15)  
  # p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.7, 0.72, 0.42) [soft-clipping]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.85, 0.35)       #old: g = c(0.2, 0.2, 0.25) 
  # w = c(0.72, 0.15, 0.62)  
# d = c(0.05, 0.7, 0.25) 

```


```{r, eval=F}
#====================================================
  # sd_y = 0.08, mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1, # sd_base = 0.2; mu_base = -1.6 
  # sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0.4, 0.15, 0.15)  
  # p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.7, 0.72, 0.42) [soft-clipping]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.85, 0.35)       #old: g = c(0.2, 0.2, 0.25) 
  # w = c(0.72, 0.15, 0.62)  
# d = c(0.05, 0.7, 0.25) 
#====================================================


mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "s[1]", "s[2]", "s[3]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "s[1]", "s[2]", "s[3]"))
mcmc_pairs(posterior, np = np, pars = c("theta", "mu_season[1]", "s[1]", "s[2]", "s[3]"))
mcmc_pairs(posterior, np = np, pars = c("mu_season[2]", "mu_season[3]", "s[1]", "s[2]", "s[3]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "s[1]", "s[2]", "s[3]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[3]", "gw_mu[4]", "s[1]", "s[2]", "s[3]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[5]", "s[1]", "s[2]", "s[3]")) #"gw_mu[6]",
mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "s[1]", "s[2]", "s[3]"))
mcmc_pairs(posterior, np = np, pars = c("p_mu[3]", "d[1]", "s[1]", "s[2]", "s[3]"))
mcmc_pairs(posterior, np = np, pars = c("d[2]", "d[3]", "s[1]", "s[2]", "s[3]"))



mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "gw_mu[2]", "gw_mu[3]"))

mcmc_pairs(posterior, np = np, pars = c("gw_mu[2]", "gw_mu[3]", "theta"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "theta"))

mcmc_pairs(posterior, np = np, pars = c("gw_mu[2]", "gw_mu[3]", "s[1]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[2]", "gw_mu[3]", "s[2]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[2]", "gw_mu[3]", "s[3]"))

mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "s[1]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "s[2]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "s[3]"))


```

```{r, eval=F} 
#====================================================
  # sd_y = 0.08, mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1, # sd_base = 0.2; mu_base = -1.6 
  # sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0.4, 0.15, 0.15)  
  # p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.7, 0.72, 0.42) [soft-clipping]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.85, 0.35)       #old: g = c(0.2, 0.2, 0.25) 
  # w = c(0.72, 0.15, 0.62)  
# d = c(0.05, 0.7, 0.25) 

mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "gw_err[3]", "gw_err[4]"))
mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "gw_sd[3]", "gw_sd[4]"))
mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "mu_u1", "sd_y"))
mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "d[1]", "d[2]", "d[3]"))

mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "gw_mu[1]", "gw_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "gw_sd[1]", "gw_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "gw_err[1]", "gw_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "gw_mu[5]", "gw_mu[6]"))
mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "gw_err[5]", "gw_err[6]"))
mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "gw_sd[5]", "gw_sd[6]"))

mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "p_mu[1]", "p_mu[2]", "p_mu[3]"))
mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "p_err[1]", "p_err[2]", "p_err[3]"))
mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "p_sd[1]", "p_sd[2]", "p_sd[3]"))

mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "ab_err[1]", "ab_err[2]"))
mcmc_pairs(posterior, np = np, pars = c("theta", "gw_mu[3]", "gw_mu[4]", "ab_sd[1]", "ab_sd[2]"))


mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "gw_mu[3]", "gw_mu[4]", "gw_mu[5]", "gw_mu[6]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "gw_mu[3]", "p_mu[1]", "p_mu[2]", "p_mu[3]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "gw_mu[3]", "ab_mu[1]", "ab_mu[2]", "sd_y"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "gw_mu[3]", "d[1]", "d[2]", "d[3]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[4]", "gw_mu[5]", "gw_mu[6]", "p_mu[1]", "p_mu[2]", "p_mu[3]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[4]", "gw_mu[5]", "gw_mu[6]", "ab_mu[1]", "ab_mu[2]", "sd_y"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[4]", "gw_mu[5]", "gw_mu[6]", "d[1]", "d[2]", "d[3]"))
mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p_mu[3]", "ab_mu[1]", "ab_mu[2]", "sd_y"))
mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p_mu[3]", "d[1]", "d[2]", "d[3]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "sd_y", "d[1]", "d[2]", "d[3]"))

mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[4]", "p_mu[1]", "d[1]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[2]", "gw_mu[5]", "p_mu[2]", "d[2]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[3]", "gw_mu[6]", "p_mu[3]", "d[3]", "ab_mu[1]", "ab_mu[2]"))

mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[4]", "p_mu[1]", "d[1]", "sd_y", "mu_u1"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[2]", "gw_mu[5]", "p_mu[2]", "d[2]", "sd_y", "mu_u1"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[3]", "gw_mu[6]", "p_mu[3]", "d[3]", "sd_y", "mu_u1"))


#====================================================
  # sd_y = 0.08, mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1, # sd_base = 0.2; mu_base = -1.6 
  # sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0.4, 0.15, 0.15)  
  # p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.7, 0.72, 0.42) [soft-clipping]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.85, 0.35)       #old: g = c(0.2, 0.2, 0.25) 
  # w = c(0.72, 0.15, 0.62)  
# d = c(0.05, 0.7, 0.25)  

```

```{r, eval=F}
  # sd_y = 0.08, mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1, # sd_base = 0.2; mu_base = -1.6 
  # sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0.4, 0.15, 0.15)  
  # p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.7, 0.72, 0.42) [soft-clipping]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.85, 0.35)       #old: g = c(0.2, 0.2, 0.25) 
  # w = c(0.72, 0.15, 0.62)  
# d = c(0.05, 0.7, 0.25) 
#====================================================

        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "g_sd[3]", "g_err[3]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "mu_u1", "sd_y"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "d[1]", "d[2]", "d[3]"))
        # 
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "w_mu[1]", "w_mu[2]", "w_mu[3]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "w_sd[1]", "w_sd[2]", "w_sd[3]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "w_err[1]", "w_err[2]", "w_err[3]"))
        # 
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "p_mu[1]", "p_mu[2]", "p_mu[3]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "p_sd[1]", "p_sd[2]", "p_sd[3]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "p_err[1]", "p_err[2]", "p_err[3]"))
        # 
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "g_mu[1]", "g_mu[2]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "g_sd[1]", "g_sd[2]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "g_err[1]", "g_err[2]"))
        # 
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "ab_mu[1]", "ab_mu[2]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "ab_sd[1]", "ab_sd[2]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "ab_err[1]", "ab_err[2]"))
        # 
        # ###
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "g_err[1]", "g_err[2]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[2]", "g_mu[3]", "g_err[2]", "g_err[3]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[3]", "g_err[1]", "g_err[3]"))
        # 
        # mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "w_err[1]", "w_err[2]"))
        # mcmc_pairs(posterior, np = np, pars = c("w_mu[2]", "w_mu[3]", "w_err[2]", "w_err[3]"))
        # mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[3]", "w_err[1]", "w_err[3]"))
        # 
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "w_mu[1]", "g_err[1]", "w_err[1]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[2]", "w_mu[2]", "g_err[2]", "w_err[2]"))
        # mcmc_pairs(posterior, np = np, pars = c("g_mu[3]", "w_mu[3]", "g_err[3]", "w_err[3]"))


mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "gw_err[1]", "gw_sd[1]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "gw_err[2]", "gw_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "gw_err[3]", "gw_sd[3]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "gw_err[4]", "gw_sd[4]"))
#mcmc_pairs(posterior, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "gw_err[5]", "gw_sd[5]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[3]", "gw_mu[4]", "gw_err[1]", "gw_sd[1]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[3]", "gw_mu[4]", "gw_err[2]", "gw_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[3]", "gw_mu[4]", "gw_err[3]", "gw_sd[3]"))
mcmc_pairs(posterior, np = np, pars = c("gw_mu[3]", "gw_mu[4]", "gw_err[4]", "gw_sd[4]"))
#mcmc_pairs(posterior, np = np, pars = c("gw_mu[3]", "gw_mu[4]", "gw_mu[5]", "gw_err[5]", "gw_sd[5]"))

mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p_mu[3]", "gw_mu[1]", "gw_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p_mu[3]",  "gw_mu[3]", "gw_mu[4]"))#, "gw_mu[5]"))
#mcmc_pairs(posterior, np = np, pars = c("p_mu[3]", "gw_mu[3]", "gw_mu[4]"))#, "gw_mu[5]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "gw_mu[1]", "gw_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "gw_mu[3]", "gw_mu[4]"))#, "gw_mu[5]"))

mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "gw_mu[1]", "gw_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "gw_mu[3]", "gw_mu[4]"))#, "gw_mu[5]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "p_mu[1]", "p_mu[2]", "p_mu[3]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "d[1]", "d[2]", "d[3]"))

mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "gw_mu[1]", "gw_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "gw_mu[3]", "gw_mu[4]"))#, "gw_mu[5]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "p_mu[1]", "p_mu[2]", "p_mu[3]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "d[1]", "d[2]", "d[3]"))


#====================================================
  # sd_y = 0.08, mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1, # sd_base = 0.2; mu_base = -1.6 
  # sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0.4, 0.15, 0.15)  
  # p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.7, 0.72, 0.42) [soft-clipping]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.85, 0.35)       #old: g = c(0.2, 0.2, 0.25) 
  # w = c(0.72, 0.15, 0.62)  
# d = c(0.05, 0.7, 0.25) 

```


# g
```{r, eval=F}
mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "g_sd[1]", "g_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "g_err[1]", "g_err[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_sd[1]", "g_sd[2]", "g_err[1]", "g_err[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "g_L[2,1]", "g_L[2,2]"))
mcmc_pairs(posterior, np = np, pars = c("g_sd[1]", "g_sd[2]", "g_L[2,1]", "g_L[2,2]"))
mcmc_pairs(posterior, np = np, pars = c("g_err[1]", "g_err[2]", "g_L[2,1]", "g_L[2,2]"))

mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "p_mu[1]", "p_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "p_sd[1]", "p_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "p_err[1]", "p_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("g_sd[1]", "g_sd[2]", "p_mu[1]", "p_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_sd[1]", "g_sd[2]", "p_sd[1]", "p_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_sd[1]", "g_sd[2]", "p_err[1]", "p_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("g_err[1]", "g_err[2]", "p_mu[1]", "p_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_err[1]", "g_err[2]", "p_sd[1]", "p_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_err[1]", "g_err[2]", "p_err[1]", "p_err[2]"))


mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("g_sd[1]", "g_sd[2]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_sd[1]", "g_sd[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_sd[1]", "g_sd[2]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("g_err[1]", "g_err[2]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_err[1]", "g_err[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("g_err[1]", "g_err[2]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "g_mu[1]", "g_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "g_sd[1]", "g_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "g_err[1]", "g_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "g_mu[1]", "g_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "g_sd[1]", "g_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "g_err[1]", "g_err[2]"))


#====================================================
  # mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1,  
  # sd_base = 0.2; mu_base = -1.6 
  # sd_y = 0.08, sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0, 0.7, 0.5)  #old: s = c(0.4, 0.7, 0.5) 
  # p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.7, 0.72, 0.42) [soft-clipping]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.3, 0.2, 0.4) 
  # w = c(0.72, 0.15, 0.62) 
# d = c(0.05, 0.7, 0.25) 

```


# w
```{r, eval=F} 
mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "w_sd[1]", "w_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "w_err[1]", "w_err[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_sd[1]", "w_sd[2]", "w_err[1]", "w_err[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "w_L[2,1]", "w_L[2,2]"))
mcmc_pairs(posterior, np = np, pars = c("w_sd[1]", "w_sd[2]", "w_L[2,1]", "w_L[2,2]"))
mcmc_pairs(posterior, np = np, pars = c("w_err[1]", "w_err[2]", "w_L[2,1]", "w_L[2,2]"))

mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "p_mu[1]", "p_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "p_sd[1]", "p_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "p_err[1]", "p_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("w_sd[1]", "w_sd[2]", "p_mu[1]", "p_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_sd[1]", "w_sd[2]", "p_sd[1]", "p_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_sd[1]", "w_sd[2]", "p_err[1]", "p_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("w_err[1]", "w_err[2]", "p_mu[1]", "p_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_err[1]", "w_err[2]", "p_sd[1]", "p_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_err[1]", "w_err[2]", "p_err[1]", "p_err[2]"))


mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("w_sd[1]", "w_sd[2]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_sd[1]", "w_sd[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_sd[1]", "w_sd[2]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("w_err[1]", "w_err[2]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_err[1]", "w_err[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("w_err[1]", "w_err[2]", "ab_err[1]", "ab_err[2]"))
#mcmc_pairs(posterior, np = np, pars = c("w_err[1]", "w_err[2]", "sd_base"))


mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "w_mu[1]", "w_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "w_sd[1]", "w_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "w_err[1]", "w_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "w_mu[1]", "w_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "w_sd[1]", "w_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "w_err[1]", "w_err[2]"))

      # mcmc_pairs(posterior, np = np, pars = c("w[1]", "w[2]", "ab_mu[1]", "ab_mu[2]"))
      # mcmc_pairs(posterior, np = np, pars = c("w[1]", "w[2]", "ab_sd[1]", "ab_sd[2]"))
      # mcmc_pairs(posterior, np = np, pars = c("w[1]", "w[2]", "ab_err[1]", "ab_err[2]"))
      # mcmc_pairs(posterior, np = np, pars = c("w[1]", "w[2]", "p_mu[1]", "p_mu[2]", "p0[1]"))
      # mcmc_pairs(posterior, np = np, pars = c("w[1]", "w[2]", "p_sd[1]", "p_sd[2]"))
      # mcmc_pairs(posterior, np = np, pars = c("w[1]", "w[2]", "p_err[1]", "p_err[2]"))
      # mcmc_pairs(posterior, np = np, pars = c("w[1]", "w[2]", "mu_u1", "sd_y"))

#====================================================
  # mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1,  
  # sd_base = 0.2; mu_base = -1.6 
  # sd_y = 0.08, sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0, 0.7, 0.5)  #old: s = c(0.4, 0.7, 0.5) 
  # p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.7, 0.72, 0.42) [soft-clipping]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.3, 0.2, 0.4) 
  # w = c(0.72, 0.15, 0.62) 
# d = c(0.05, 0.7, 0.25) 

```



# More pairs plot
```{r, eval=F}
mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p0[1]", "p_sd[1]", "p_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p0[1]", "p_err[1]", "p_err[2]"))
mcmc_pairs(posterior, np = np, pars = c("p_sd[1]", "p_sd[2]", "p_err[1]", "p_err[2]"))
# mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p0[1]", "p_L[2,1]", "p_L[2,2]"))
# mcmc_pairs(posterior, np = np, pars = c("p_sd[1]", "p_sd[2]", "p_L[2,1]", "p_L[2,2]"))
# mcmc_pairs(posterior, np = np, pars = c("p_err[1]", "p_err[2]", "p_L[2,1]", "p_L[2,2]"))

mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p0[1]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p0[1]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p0[1]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("p_sd[1]", "p_sd[2]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("p_sd[1]", "p_sd[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("p_sd[1]", "p_sd[2]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("p_err[1]", "p_err[2]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("p_err[1]", "p_err[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("p_err[1]", "p_err[2]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "p_mu[1]", "p_mu[2]", "p0[1]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "p_sd[1]", "p_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "p_err[1]", "p_err[2]"))

# mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "p_L[2,1]", "p_L[2,2]"))
# mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p0[1]", "ab_L[2,1]", "ab_L[2,2]"))
# mcmc_pairs(posterior, np = np, pars = c("p_sd[1]", "p_sd[2]", "ab_L[2,1]", "ab_L[2,2]"))
# mcmc_pairs(posterior, np = np, pars = c("p_err[1]", "p_err[2]", "ab_L[2,1]", "ab_L[2,2]"))

mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "ab_err[1]", "ab_err[2]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "p_mu[1]", "p_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "p0[1]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "p_sd[1]", "p_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "p_err[1]", "p_err[2]"))
mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "mu_u1", "sd_y"))



#====================================================
  # mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1,  
  # sd_base = 0.2; mu_base = -1.6 
  # sd_y = 0.08, sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0, 0.7, 0.5)  #old: s = c(0.4, 0.7, 0.5) 
  # p = c(0.62, 0.45, 0.3) 
  # g = c(0.3, 0.2, 0.4) 
  # w = c(0.72, 0.15, 0.62) 
# d = c(0.05, 0.7, 0.25)  
```




```{r, eval=F} 
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "ab_err[1]", "ab_err[2]"))
mcmc_pairs(posterior, np = np, pars = c("ab_sd[1]", "ab_sd[2]", "ab_err[1]", "ab_err[2]"))
# mcmc_pairs(posterior, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "ab_L[2,1]", "ab_L[2,2]"))
# mcmc_pairs(posterior, np = np, pars = c("ab_sd[1]", "ab_sd[2]", "ab_L[2,1]", "ab_L[2,2]"))
# mcmc_pairs(posterior, np = np, pars = c("ab_L[2,1]", "ab_L[2,2]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "ab_sd[1]", "ab_sd[2]"))
# mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "ab_L[2,1]", "ab_L[2,2]"))
mcmc_pairs(posterior, np = np, pars = c("mu_u1", "sd_y", "ab_err[1]", "ab_err[2]"))
mcmc_pairs(posterior, np = np, pars = c("mu_alpha", "beta", "mu_u1", "sd_y")) 



            # mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "p_sd[1]", "p_sd[2]"))
            # mcmc_pairs(posterior, np = np, pars = c("p_sd[1]", "p_sd[2]", "p_err[1]", "p_err[2]"))
            # mcmc_pairs(posterior, np = np, pars = c("p_err[1]", "p_err[2]", "p_mu[1]", "p_mu[2]"))
            # 
            # mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "mu_alpha", "beta")) 
            # mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "sd_y", "mu_base")) 
            # mcmc_pairs(posterior, np = np, pars = c("p_mu[1]", "p_mu[2]", "d[1]", "d[2]", "d[3]")) 


# mcmc_pairs(posterior, np = np, pars = c("p[1]", "p[2]", "mu_alpha", "beta"))
# mcmc_pairs(posterior, np = np, pars = c("p[1]", "p[2]", "sd_y", "mu_base"))
# mcmc_pairs(posterior, np = np, pars = c("g[1]", "g[2]", "mu_alpha", "beta"))
# mcmc_pairs(posterior, np = np, pars = c("g[1]", "g[2]", "sd_y"))#, "mu_base"))
# mcmc_pairs(posterior, np = np, pars = c("w[1]", "w[2]", "mu_alpha", "beta"))
# mcmc_pairs(posterior, np = np, pars = c("w[1]", "w[2]", "sd_y"))#, "mu_base"))
# 
# mcmc_pairs(posterior, np = np, pars = c("p[1]", "p[2]", "g[1]", "g[2]"))
# mcmc_pairs(posterior, np = np, pars = c("p[1]", "p[2]", "w[1]", "w[2]"))
# mcmc_pairs(posterior, np = np, pars = c("g[1]", "g[2]", "w[1]", "w[2]"))

# mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "ab_mu[1]", "ab_mu[2]")) 
# mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "ab_sd[1]", "ab_sd[2]")) 
# mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "ab_err[1]", "ab_err[2]")) 
# mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "mu_alpha", "beta"))#, "theta"))))#, #,"w", "d"))
# mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "mu_u1", "sd_y"))#, "mu_base"))#, "theta"))))#, 


#mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "p[1]", "p[2]"))#, "theta"))))#, 
# mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "g[1]", "g[2]"))#, "theta"))))#, 
# mcmc_pairs(posterior, np = np, pars = c("d[1]", "d[2]", "d[3]", "w[1]", "w[2]"))#, "theta"))))#, 



            # mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "g_sd[1]", "g_sd[2]"))
            # mcmc_pairs(posterior, np = np, pars = c("g_sd[1]", "g_sd[2]", "g_err[1]", "g_err[2]"))
            # mcmc_pairs(posterior, np = np, pars = c("g_err[1]", "g_err[2]", "g_mu[1]", "g_mu[2]"))
            # 
            # mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "w_sd[1]", "w_sd[2]"))
            # mcmc_pairs(posterior, np = np, pars = c("w_sd[1]", "w_sd[2]", "w_err[1]", "w_err[2]"))
            # mcmc_pairs(posterior, np = np, pars = c("w_err[1]", "w_err[2]", "w_mu[1]", "w_mu[2]"))
            # 
            # mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "mu_alpha", "beta")) 
            # mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "sd_y", "mu_base")) 
            # mcmc_pairs(posterior, np = np, pars = c("g_mu[1]", "g_mu[2]", "d[1]", "d[2]", "d[3]")) 
            # 
            # mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "mu_alpha", "beta")) 
            # mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "sd_y", "mu_base")) 
            # mcmc_pairs(posterior, np = np, pars = c("w_mu[1]", "w_mu[2]", "d[1]", "d[2]", "d[3]")) 


#====================================================
  # mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1,  
  # sd_base = 0.2; mu_base = -1.6 
  # sd_y = 0.08, sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0, 0.7, 0.5)  #old: s = c(0.4, 0.7, 0.5) 
  # p = c(0.62, 0.45, 0.3) 
  # g = c(0.3, 0.2, 0.4) 
  # w = c(0.72, 0.15, 0.62) 
# d = c(0.05, 0.7, 0.25) 

```


```{r, eval=F}
#library(shinystan)
#??? ss_complete_pool <- extract(fit2);

library(shinystan)
ss_output <- launch_shinystan(as.shinystan(fit2, 
                              pars=c(
                                     "ab_mu", "ab_sd", "ab_L", "ab_L", "ab_err", 
                                     "mu_alpha", "beta", #"mu_base", #"theta", "sd_base", #"rho", 
              "sd_y", #"sd_season", "mu_season", #"s", "p", 
                 # "ab_mu", "ab_sd", "ab_L",
                 # "p_mu", "p_sd", "p_L",
                 # "g_mu", "g_sd", "g_L",
                 # "w_mu", "w_sd", "w_L",
                 # "p",
                 # "g", "w", "d",
              "mu_u1"
              )))

```





```{r, eval=F}
color_scheme_set("mix-brightblue-gray")
mcmc_trace(posterior_cp, pars = "tau", np = np_cp) +
  xlab("Post-warmup iteration")

mcmc_trace(posterior_cp, pars = "tau", np = np_cp, window = c(400,550)) +
  xlab("Post-warmup iteration")

mcmc_nuts_divergence(np_cp, lp_cp, chain = 4)
mcmc_nuts_divergence(np_cp, lp_cp)

color_scheme_set("red")
mcmc_nuts_energy(np_cp)

rhats <- rhat(fit_cp_bad_rhat)
print(rhats)
color_scheme_set("brightblue") # see help("color_scheme_set")
mcmc_rhat(rhats)
mcmc_rhat(rhats) + yaxis_text(hjust = 1)
mcmc_rhat(rhat = rhat(fit_cp)) + yaxis_text(hjust = 0)
ratios_cp <- neff_ratio(fit_cp)
print(ratios_cp)

neff_cp <- neff_ratio(fit_cp, pars = c("theta", "mu", "tau"))
neff_ncp <- neff_ratio(fit_ncp, pars = c("theta", "mu", "tau"))
compare_cp_ncp(mcmc_neff(neff_cp), mcmc_neff(neff_ncp), ncol = 1)

compare_cp_ncp(
  mcmc_acf(posterior_cp, pars = "theta[1]", lags = 10),
  mcmc_acf(posterior_ncp, pars = "eta[1]", lags = 10)
)

```


```{r, eval=F}
# pairs_stan(1, fit2, c("rho_LT", "sd_pi", "mu_u1", "mu_alpha", "beta",  "mu_season"))
# pairs_stan(2, fit2, c("rho_LT", "sd_pi", "mu_u1", "mu_alpha", "beta",  "mu_season"))
# pairs_stan(1, fit2, c("rho_LT", "sd_pi", "sd_gamma", "sd_omega", "sd_season", "sd_base", "sd_y"))
# pairs_stan(2, fit2, c("rho_LT", "sd_pi", "sd_gamma", "sd_omega", "sd_season", "sd_base", "sd_y"))
# pairs_stan(1, fit2, c("rho_LT", "sd_pi", "p", "w"))
# pairs_stan(2, fit2, c("rho_LT", "sd_pi", "p", "w"))
# pairs_stan(1, fit2, c("rho_LT", "sd_pi", "p", "g"))
# pairs_stan(2, fit2, c("rho_LT", "sd_pi", "p", "g"))
# pairs_stan(1, fit2, c("rho_LT", "sd_pi", "p", "d"))
# pairs_stan(2, fit2, c("rho_LT", "sd_pi", "p", "d"))

# pairs_stan(1, fit2, c("sd_season", "mu_season"))
# pairs_stan(2, fit2, c("sd_season", "mu_season"))
# pairs_stan(1, fit2, c("rho_LT", "sd_pi", "p"))  
# pairs_stan(2, fit2, c("rho_LT", "sd_pi", "p"))  
pairs_stan(1, fit2, c("mu_u1", "mu_alpha", "beta", "sd_y", "sd_season", "mu_season"))  
pairs_stan(1, fit2, c("mu_alpha", "beta", #"p"
                      "g", 
                      "w", "d"
                      ))#"season_raw"))#, "sd_pi", "eta", "s", 


#pairs_stan(2, fit2, c("mu_u1", "mu_alpha", "beta", "sd_y"))#, "sd_base", "sd_gamma", "sd_omega"))
# pairs_stan(1, fit2, c("sd_base", "g", "w", "d"))
# pairs_stan(2, fit2, c("sd_base", "g", "w", "d"))


#pairs(fit, pars = c("A", "lp__","z[4]")) 


# traceplot of the problematic parameter
params.df <- as.data.frame(extract(fit2, permuted=FALSE))
names(params.df) <- gsub("chain:1.", "", names(params.df), fixed = TRUE)
names(params.df) <- gsub("[", ".", names(params.df), fixed = TRUE)
names(params.df) <- gsub("]", "", names(params.df), fixed = TRUE)
params.df$iter <- 1:(n_iter-n_warmup)

par(mar = c(4, 4, 0.5, 0.5))
plot(params.df$iter, params.df$mu_u1, pch=16, cex=0.8, #col=c_dark, 
     xlab="Iteration")#, ylab="log(tau)", ylim=c(-6, 4))
plot(params.df$iter, params.df$mu_alpha, pch=16, cex=0.8, #col=c_dark, 
     xlab="Iteration")#, ylim=c(0, 0.3))#, ylab="log(tau)", ylim=c(-6, 4))
plot(params.df$iter, params.df$beta, pch=16, cex=0.8, #col=c_dark, 
     xlab="Iteration")#, ylim=c(0, 0.3))#, ylab="log(tau)", ylim=c(-6, 4))
plot(params.df$iter, params.df$sd_y, pch=16, cex=0.8, #col=c_dark, 
     xlab="Iteration")#, ylab="log(tau)", ylim=c(-6, 4))
plot(params.df$iter, params.df$sd_season, pch=16, cex=0.8, #col=c_dark, 
      xlab="Iteration")#, ylab="log(tau)", ylim=c(-6, 4))
#plot(params.df$iter, params.df$eta, pch=16, cex=0.8, #col=c_dark, 
#      xlab="Iteration")#, ylab="log(rho_raw)")#, ylim=c(-6, 4))


# plot(params.df$iter, params.df$sd_gamma, pch=16, cex=0.8, #col=c_dark, 
#      xlab="Iteration")#, ylim=c(0, 0.15))#, ylab="log(tau)", 
# plot(params.df$iter, params.df$sd_omega, pch=16, cex=0.8, #col=c_dark, 
#      xlab="Iteration")#, ylim=c(0, 0.3))#, ylab="log(tau)", ylim=c(-6, 4))
# plot(params.df$iter, params.df$sd_base, pch=16, cex=0.8, #col=c_dark, 
#      xlab="Iteration")#, ylim=c(0, 0.2))#, ylab="log(tau)", ylim=c(-6, 4))



    # traceplot(fit,
    #           c("Omega", "alpha", "beta", "sigma_uv", "rho_uv",
    #             "uv[1,1]", "uv[1,2]", "uv[30,1]", "uv[30,2]",
    #             "lp__"),
    #           inc_warmup=FALSE) +
    #   coord_cartesian(xlim = c(3001,3200)) +
    #   scale_x_continuous(breaks=c(3000, 3100, 3200))



```


```{r, eval = F}

# # extract base that limits the lowest P reachable
# m.mat %>% t() %>% psych::describe() %>% print(digits=3)
# D.mat %>% t() %>% psych::describe() %>% print(digits=3)
# y.mat %>% t() %>% psych::describe() %>% print(digits=3)
# { base.mat <- extract(simu_fit)$base[1,] } %>% t()
# psych::describe(base.mat) %>% print(digits=3)
# extract(simu_fit)$p_mu[1,,] %>% t() %>% psych::describe() %>% print(digits=3)
# extract(simu_fit)$b[1,,] %>% t() %>% psych::describe() %>% print(digits=3)

#//===============================================================================
      # m.mat <- extract(simu_fit)$m[1,,]
      # 
      # if (is.vector(m.mat)) m.mat <- matrix(m.mat, nrow = 1)
      # D.mat <- apply(m.mat, MARGIN=1, 
      #                function(m) {  # margin=1 means one after another by row   # , L, d)
      #                 # D <- m - d[1]*lag(m) - d[2]*lag(m,2L) - d[3]*lag(m,3L)
      #                 # D[1] = m[1]
      #                 # D[2] = m[2] - d[1]*m[1]     
      #                 # D[3] = m[3] - d[1]*m[2] - d[2]*m[1]
      #                  D <- m
      #                  for (i in 1:L) { D <- (D - d[i]*dplyr::lag(m,i, default = 0)) }
      #                  D
      #                  }
      #                ) %>%   
      #             t()
#//===============================================================================

#r.mat <- (y.mat) %>% matrix(nrow=J, ncol=N) 

#//===============================================================================
#r.l <- (y.mat) %>%  
    #r.l <- (y.mat + D.mat) %>%  
#  t() %>% 
#  split(col(.)) %>% 
#  lapply(as.vector)
#//===============================================================================



# For debugging only
      # RealST.l <- extract(simu_fit)$RealST[1,,] %>%
      #   t() %>%
      #   split(col(.)) %>%
      #   lapply(as.vector)
      # RealST.l
      # lapply(RealST.l, mean) %>% unlist() %>%  mean()



#The higher lambda is, the more benefit pooling there is and the greater the benefit from [non-centered] parameterization in hierarchical 
models
# https://groups.google.com/d/msg/stan-users/HWiUtQLRCfE/RXT17CFcDQAJ
lambda <- function(fit, ...) {
    extracted_fit <- rstan::extract(fit, permuted = TRUE, ...)
    N <- length(extracted_fit)
    result <- rep(NA, N)
    names(result) <- names(extracted_fit)
    for (i in 1:N) {
        extracted_fit_i <- extracted_fit[[i]]
        if (length(dim(extracted_fit_i)) == 1) next #only calculate if more than 
                                                    #1 dimension
        e <- extracted_fit_i - mean(extracted_fit_i)
        result[i] <- 1 - var(apply(e, 2, mean)) / mean(apply(e, 1, var))
    }
    return(result)
}

#lambda(fit1)

warnings()
```

```{r}
sessionInfo()
```


