---
title: "State space model of accounting manipulation - Stan"
output: html_notebook
---

```{r}

#library(tidyverse)
library(dplyr)
library(magrittr)

#load previously saved Stan model objects
#load(file = "StanObjects.rda")

library(rstan)
rstan_options(auto_write = TRUE) # avoid recompilation of unchanged Stan programs

# The following can affect distributional computing over a cluster
options(mc.cores = parallel::detectCores())  # 

# The following throws an error in compiling .stan
#Sys.setenv(LOCAL_CPPFLAGS = '-march=native') # For improved execution time but can throw errors for some CPUs

```


# Estimate the simulated model with Stan - misreporting extent 
  (forecasts specified in generated quantities block)
//ignore these lines: 
//  COGS2Rev, SGA2Rev, RnD2Rev with Rev2Rev = unit vector as the intercept)
//   int<lower=0,upper=1> M[N];   // M = 1 for the decision to misreport; = 0 if report honestly  

```{stan output.var=SSM, eval=F}
```

# Simulate data
```{stan output.var=simu_SSM, eval=F}
```


# Simulate data and fit the debug model (also, save Stan model objects) 
```{r}

J = 5#5#15 #00#50    # number of firms
N = 16*2#4#2 #4 #7 number of (yearly) observations of a firm


#sd_c_init = 0.1  #0.01
sd_gamma = 0.02
sd_omega = 0.5
sd_y = 0.08  # sd of underlying unbiased figures is likely to be industry-specific / business model specific
#sd_m = 0.02  # sd of misreporting extent m, which should be restrained by the governance setting (eg, audit committee chair with financial expertise, big 4 auditor, etc)

alpha = 0.04  # intercept parameter of the AR(1) processs of the underlying unbiased figure
beta = 0.6  # slope parameter of the AR(1) processs of the underlying unbiased figure
#set parameters of the data generation process of the misreporting extent m
#g =c(-0.07, 0.2)
g = c(0.05, -0.1, 0.2)
K = length(g)
#cvec <- matrix(data = g, nrow = K)  # define column vector as a matrix[N,1]
w = c(9, 4, 0.7)
H = length(w)
d = c(0.8, 0.2) # fractional reversal of prior-period manipluation by accruals
#d = c(0.6, 0.3, 0.1) # fractional reversal of prior-period manipluation by accruals
L = length(d)

ratioNED_req = 0.25  # min ratio of non-executive directors (NED) required by law
integrity = 0.5 #(0,1)  0.5 = average       
# integrity level, between (0,1), of the society, interpreted as likelihood to behave opportunistically
# integrity level of an individual, indiv_integ between (-Inf,Inf), is 
#   affected by a society's general integrity level in a stochastic fashion varying from period to period

#M <- rbinom(N, 1, integrity)   # the decision to misreport earnings


# Define function to convert data.frame with gvkey as the key to a list of matices
df2mat.l <- function(z){
  z %>% 
  split(.$gvkey) %>% 
  lapply(function(z) select(z, -c(gvkey))) %>% 
  lapply(as.matrix)
  }

#set.seed(8)
# covariate matrix X capturing the dis/incentive to misreport
# downgrade = 1 if one or more analysts following the firm downgraded their stock recommendation in the last period
X.l <- data.frame(
  gvkey = rep(1:J, times=N),                   
  constant = rep(1, J*N),
  indiv_integ = rbeta(J*N, integrity, 1-integrity), 
  downgrade = rbinom(J*N, 1, 0.3)
  ) %>%
  df2mat.l()
  # split(.$gvkey) %>% 
  # lapply(function(z) select(z, -c(gvkey))) %>% 
  # lapply(as.matrix)

# covariate matrix G capturing the strength of internal governance and external monitoring mechansims
G.l <- data.frame(
  gvkey = rep(1:J, each=N), # note the difference from times=N
  constant = rep(1, J*N),
  ratioNED = rep(rbeta(J, ratioNED_req, 1-ratioNED_req), each=N),
  Big4 = rbinom(J*N, 1, 0.7)
  ) %>%
  df2mat.l()

#colnames(Z) <- c("Rev2Rev", "COGS2Rev", "SGA2Rev", "RnD2Rev")

simu_SSM_data <- list(J = J, 
                      N = N, 
                      K = K,
                      H = H,
                      X = X.l,
                      G = G.l,
                      alpha = alpha,
                      beta = beta, 
                      g = g,
                      w = w,
                      sd_gamma = sd_gamma,
                      sd_omega = sd_omega,
                      sd_y = sd_y#,
#                      sd_m = sd_m
                      )
# save(simu_SSM_data, file="simu_SSM_data.rda")
# load(file="simu_SSM_data.rda")

# Simulate SSM data
simu_fit <- stan(#model_code = simu_SSM@model_code,
                 file = "simu_SSM1.7.stan",  # Stan program
                 data=simu_SSM_data, iter=1, #seed=987,  
                 chains=1, algorithm="Fixed_param")  

y.mat <- extract(simu_fit)$y[1,,]
m.mat <- extract(simu_fit)$m[1,,]

D.mat <- apply(m.mat, MARGIN=1, 
               function(m) {  # margin=1 means one after another by row   # , L, d)
                # D <- m - d[1]*lag(m) - d[2]*lag(m,2L) - d[3]*lag(m,3L)
                # D[1] = m[1]
                # D[2] = m[2] - d[1]*m[1]     
                # D[3] = m[3] - d[1]*m[2] - d[2]*m[1]
                 D <- m
                 for (i in 1:L) { D <- (D - d[i]*dplyr::lag(m,i, default = 0)) }
                 D
                 }
               ) %>%   
            t()

r.l <- (y.mat + D.mat) %>%  
  t() %>% 
  split(col(.)) %>% 
  lapply(as.vector)


# Prepare SSM data for estimation using MCMC
#N_new = 3
SSM_data <- list(J = J,
                 N = N, 
                 K = K,
                 H = H,
                 L = L,
                 sd_gamma_init = sd_gamma,
                 sd_omega_init = sd_omega,
                 sd_y_init = sd_y,
#                 sd_m_init = sd_m,
                 alpha_init = alpha,
                 beta_init = beta,
                 g_init = g,
                 w_init = w,
#                 sd_c_init = sd_c_init,
                 r = r.l, 
                 X = X.l,
                 G = G.l
                 # Z_new = Z[1:N_new,],
                 # N_new = N_new
                 )
# Fit the debug model to make sure the Stan code complies properly 
#fit0_debug <- stan(model_code = SSM@model_code, data = SSM_data, iter = 10, chains = 1)

#save(simu_GP, misreport, file = "StanObjects.rda")
#load(file = "StanObjects.rda")
```

# Fit the full model (by referring to the debug model) to estimate underlying parameters
```{r}

# Run the full model and refer to the debug model to save compilation time 
system.time({
fit1 <- stan(
  file = "SSM1.7.stan",  # Stan program
  data = SSM_data,    # named list of data
#  model_code = SSM@model_code, 
#  fit = fit0_debug,   # to save compilation time if the debug model was run
  control = list(adapt_delta = 0.999, 
#                 stepsize = 0.05, #0.05, #0.01
                 max_treedepth = 10),    # adjust when there're divergent transitions after warmup
#  chains = 1,#8,             # default is 4 Markov chains
#  cores = 8,
#  init = init, # where init = list(list(mu=...,sigma=...), list(mu=..., sigma=...), ...) 
                # where length of list = number of chains
#  seed = 555, 

#  iter = 6000,#4000,#7000,#12000, #500 #2000,            # total number of iterations per chain
#  warmup = 4000,#500,          # number of warmup iterations per chain
  refresh = 1000#0          # = 0 means no progress shown
  )
})

beepr::beep()

#extract(fit1)$y[1,]
#pairs(extract(fit1), pars = c(a, b)) #, y[1], sd_y, m[1], sd_m, c[1], c[2], c[3]))

```

# Summary of the posterior sample 
```{r}

ss_complete_pool <- extract(fit1);
# rho_T <- 5.5, alpha_T <- 3, sigma_T <- 2
print(fit1, probs = c(0.1, 0.5, 0.9), digits = 3) #, probs = c(.05,.95))

# source('stan_utility.R')
# check_all_diagnostics(fit1)

    # alpha = 0.04  # intercept parameter of the AR(1) processs of the underlying unbiased figure
    # beta = 0.6  # slope parameter of the AR(1) processs of the underlying unbiased figure
    # g = c(0.05, -0.1, 0.2)
    # w = c(9, 4, 0.7)
    # d = c(0.6, 0.3, 0.1) # fractional reversal of prior-period manipluation by accruals
    # sd_y = 0.08  # sd of underlying unbiased figures is likely to be industry-specific/bus-model specific
    # sd_gamma = 0.02
    # sd_omega = 1

```

```{r}

#The higher lambda is, the more benefit pooling there is and the greater the benefit from [non-centered] parameterization in hierarchical models
# https://groups.google.com/d/msg/stan-users/HWiUtQLRCfE/RXT17CFcDQAJ
lambda <- function(fit, ...) {
    extracted_fit <- rstan::extract(fit, permuted = TRUE, ...)
    N <- length(extracted_fit)
    result <- rep(NA, N)
    names(result) <- names(extracted_fit)
    for (i in 1:N) {
        extracted_fit_i <- extracted_fit[[i]]
        if (length(dim(extracted_fit_i)) == 1) next #only calculate if more than 
                                                    #1 dimension
        e <- extracted_fit_i - mean(extracted_fit_i)
        result[i] <- 1 - var(apply(e, 2, mean)) / mean(apply(e, 1, var))
    }
    return(result)
}

#lambda(fit1)

warnings()
sessionInfo()
```



