---
title: "tidyk"
output: html_notebook
---

 
# Load packages
```{r}
library(magrittr)
library(dplyr)
library(forcats)
library(tidyr)
library(modelr)
library(tidybayes)
library(ggplot2)
library(ggstance)
library(ggridges)
library(cowplot)
library(rstan)
library(brms)
library(ggrepel)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

# Generate data: 10 individuals, each has different responses to the 5 conditions (A-E)
```{r}

set.seed(5)
n = 10
n_condition = 5
ABC =
  data_frame(
    condition = rep(c("A","B","C","D","E"), n),
    response = rnorm(n * 5, c(0,1,2,1,-1), 0.5)
  )

# A snapshot of the data
head(ABC, 10)

```

# Fit a hierarchical model, response ~ (1|condition), with shrinkage towards a *global mean* for each condition
```{r}

m <- brm(response ~ (1|condition), data = ABC, control = list(adapt_delta = .99),
  prior = c(
    prior(normal(0, 1), class = Intercept),  # global mean, so only one coefficient 
    prior(student_t(3, 0, 1), class = sd),    # ? sd of the est'd intercept estimate (follows student t?)
    prior(student_t(3, 0, 1), class = sigma)   # what's signma? 
  ))

m

get_variables(m) 
# b_Intercept is the global mean 
# r_condition[] variables are offsets from that mean for each condition

```

# Example: provide spread_draws with a column specification [condition,term] for variable r_condition[,]
```{r}

m %>%
  spread_draws(r_condition[condition,term]) %>%
#  spread_draws(r_condition[condition,]) %>% # Could have omitted 'term' as there's no variation in this model
  View() #  head(10)

# Note: If you have used spread_draws with a raw sample from Stan or JAGS, you may be used to using recover_types before spread_draws to get index column values back (e.g. if the index was a factor). This is not necessary when using spread_draws on rstanarm models, because those models already contain that information in their variable names. For more on recover_types, see vignette(“tidybayes”).

```

# family of functions for generating point summaries and intervals from draws: naming scheme [median|mean|mode]_[qi|hdi]
```{r}
# qi yields a quantile interval (a.k.a. equi-tailed interval, central interval, or percentile interval)  
# hdi yields a highest (posterior) density interval. 
# Custom point summary or interval functions can also be applied using the point_interval function.

m %>%
  spread_draws(b_Intercept, sigma) %>%  # gives a tidy data frame for b_Intercept and sigma
  median_qi()  # use every column not a grouping column / special column (like .chain, .iteration, or .draw)

# Use long-format 
m %>%
  gather_draws(b_Intercept, sigma) %>%  # gives a tidy data frame for b_Intercept and sigma
  median_qi()  # use only b_Intercept column

```

# The following 'group_by' operations are the same
```{r}

m %>%
  spread_draws(r_condition[condition,]) %>%
  group_by(condition) %>%   # this line not necessary (done by spread_draws)
  median_qi(r_condition)


m %>%
  spread_draws(r_condition[condition,]) %>%
  median_qi()

```

# Compute/Plot the mean within each condition, ie, b_Intercept + the effect for a given condition (r_condition)
```{r}

# spread_draws and gather_draws support extracting variables that have different indices into the same data frame. Indices with the same name are automatically matched up, and values are duplicated as necessary to produce one row per all combination of levels of all indices

m %>%
  spread_draws(`b_Intercept`, r_condition[condition,]) %>% # show columns automatching by the condition column
  mutate(condition_mean = b_Intercept + r_condition) %>%
#  View() 
  median_qi(condition_mean)

# can simplify to the below, followed by a plot of the mean within each condition:
m %>%
  spread_draws(b_Intercept, r_condition[condition,]) %>%
  median_qi(condition_mean = b_Intercept + r_condition) %>% 
# plot the mean for each condition
  ggplot(aes(y = condition, x = condition_mean, xmin = .lower, xmax = .upper)) +
  geom_pointrangeh()

# This time with overlaid additional interval:
m %>%
  spread_draws(b_Intercept, r_condition[condition,]) %>%
  median_qi(condition_mean = b_Intercept + r_condition, .width = c(.95, .66)) %>%
  ggplot(aes(y = condition, x = condition_mean)) +
  geom_pointintervalh() # <---- not geom_pointrangeh()

# Eye plot to also show the density:
m %>%
  spread_draws(b_Intercept, r_condition[condition,]) %>%
  mutate(condition_mean = b_Intercept + r_condition) %>%
  ggplot(aes(y = condition, x = condition_mean)) +
  geom_halfeyeh() 


```

# Quantile dotplot and Posterior predictions
```{r}

## Shorthand for the operations of the previous code chunk
ABC %>%
  data_grid(condition) %>%
  add_fitted_draws(m) %>%
  ggplot(aes(x = .value, y = condition)) +
  stat_pointintervalh(.width = c(.66, .95))

# Quantile dotplot as an alternative
ABC %>%
  data_grid(condition) %>%
  add_fitted_draws(m) %>%
  do(data_frame(.value = quantile(.$.value, ppoints(100)))) %>%
  ggplot(aes(x = .value)) +
  geom_dotplot(binwidth = .04) +
  facet_grid(fct_rev(condition) ~ .) +
  scale_y_continuous(breaks = NULL)

# Posterior predictions:
# Where add_fitted_draws is analogous to brms::fitted.brmsfit (or brms::posterior_linpred), 
# add_predicted_draws is analogous to brms::predict.brmsfit (brms::posterior_predict), giving draws from the posterior predictive distribution.
ABC %>%
  data_grid(condition) %>%
  add_predicted_draws(m) %>%
  ggplot(aes(x = .prediction, y = condition)) +
  geom_density_ridges()

```

# Stan model for showing 'compose_data()'
```{stan output.var= ABC_stan}

// This model expects these variables as input:
//
// A hierarchical model of our example data might fit: 
//
// an overall mean across the conditions (overall_mean), 
// the standard deviation of the condition means (condition_mean_sd), 
// the mean within each condition (condition_mean[condition]) and 
// the standard deviation of the responses given a condition mean (response_sd):

// n: number of observations
//n_condition: number of conditions
//condition: a vector of integers indicating the condition of each observation
//response: a vector of observations

data {
  int<lower=1> n;
  int<lower=1> n_condition;
  int<lower=1, upper=n_condition> condition[n];
  real response[n];
}
parameters {
  real overall_mean;
  vector[n_condition] condition_zoffset;
  real<lower=0> response_sd;
  real<lower=0> condition_mean_sd;
}
transformed parameters {
  vector[n_condition] condition_mean;
  condition_mean = overall_mean + condition_zoffset * condition_mean_sd;
}
model {
  response_sd ~ cauchy(0, 1);         // => half-cauchy(0, 1)
  condition_mean_sd ~ cauchy(0, 1);   // => half-cauchy(0, 1)
  overall_mean ~ normal(0, 5);
  condition_zoffset ~ normal(0, 1);   // => condition_mean ~ normal(overall_mean, condition_mean_sd)
  for (i in 1:n) {
    response[i] ~ normal(condition_mean[condition[i]], response_sd);
  }
}

```

# Using compose_data()
```{r}

# compose_data() can generate a list containing the above variables in the correct format automatically.

m <- sampling(object = ABC_stan, data = compose_data(ABC), control = list(adapt_delta=0.99))

```

# View stan results and see them in tidy format
```{r}

print(m, pars = c("overall_mean", "condition_mean_sd", "condition_mean", "response_sd"))

# recover the missing type information by passing the model through recover_types
m %<>% recover_types(ABC)

m %>%
  spread_draws(overall_mean, response_sd) %>%
  head(10)

```

# Plot
```{r}
m %>%
  spread_draws(condition_mean[condition]) %>%
  ggplot(aes(y = fct_rev(condition), x = condition_mean)) +
  stat_pointintervalh()
```

