---
title: "Traceplot for CmdStan output"
output: 
  html_notebook:
    fig_height: 7
    fig_width: 7
---


```{r}

#library(tidyverse)
library(dplyr)
library(magrittr)
library(rstan)
library(bayesplot)

# stan.tb <- read_csv("df_samples1.csv")
# myStan.tb <- stan.tb %>% 
#   select(-contains("err_y."), -contains("m."), -contains("Real."),
#          -contains("season_raw."), -contains("season_n."), -contains("season_q.")) %>% 
#   select(mu_alpha, beta, mu_u1, theta, sd_y, 
#          starts_with("p."), starts_with("g."), starts_with("w."), starts_with("d."),
#          sd_season, starts_with("mu_season."),
#          divergent__, treedepth__, energy__, 
#          everything())
# colnames(myStan.tb)

#=================================
metric = "dense_"#"diag_"#
#=================================
csvfiles <- dir("D:/github/learn_Stan",#system.file('misc', package = 'rstan'), 
#                pattern = '^samples[1-4].csv', full.names = TRUE)
                pattern = paste0("^", metric, "samples[1-8].csv"), full.names = TRUE)
fit <- read_stan_csv(csvfiles)

# Prepare objects for bayesplot::
fit.ar <- as.array(fit)
lp <- log_posterior(fit)
np <- nuts_params(fit)  # nuts parameters

```

# Summary of the posterior sample 
```{r} 
#====================================================
  # sd_y = 0.08, mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1, # sd_base = 0.2; mu_base = -1.6 
  # sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0.4, 0.15, 0.15)  
  # p = c(0.7, 0.62, 0.68) ; p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.85, 0.35)       #old: g = c(0.2, 0.2, 0.25) 
  # w = c(0.72, 0.15, 0.62)  
# d = c(0.05, 0.7, 0.25) 
#====================================================

print(fit, c(
              "sd_y", "mu_u1", "mu_alpha", "beta", #"rho", 
              "theta", "sd_season", "mu_season", #"sd_base", #"mu_base", 
#              "s",
              "p",
              "g", 
              "w", 
              "d", 
              "gw_mu", "gw_sd", "gw_L", "gw_err", 
              "p_mu", "p_sd", "p_L", "p_err", 
              "ab_mu", "ab_sd", "ab_L", "ab_err"
              ), 
      probs = c(0.05, 0.5, 0.95), digits = 3) #, probs = c(.05,.95))

cat("\nn_leapfrog:\n")
get_num_leapfrog_per_iteration(fit) %>% summary()
cat("\nelapsed_time:\n")
get_elapsed_time(fit)
cat("\nsampler_params chain[[1]]:\n")
get_sampler_params(fit)[[1]] %>% summary()
cat("\nseed:\n")
get_seed(fit)
cat("\n")

#====================================================
  # sd_y = 0.08, mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1, # sd_base = 0.2; mu_base = -1.6 
  # sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0.4, 0.15, 0.15)  
  # p = c(0.7, 0.62, 0.68); p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.85, 0.35)       #old: g = c(0.2, 0.2, 0.25) 
  # w = c(0.72, 0.15, 0.62)  
# d = c(0.05, 0.7, 0.25) 
#====================================================

```


```{r, fig.width=11.7, fig.height=8.3}
#```{r}


color_scheme_set("brightblue") # see help("color_scheme_set")
energy <- nuts_params(fit) %>% 
   mcmc_nuts_energy(np)  

energy + labs(
    title = paste0("Energy plot - ", metric) 
#    title = "50% posterior predictive intervals \nvs observed miles per gallon",
#    subtitle = "by vehicle weight"
  )
# 
# mcmc_nuts_divergence(np, lp)
 
# help("color_scheme_set") 
color_scheme_set("viridis")    
mcmc_trace(fit.ar, pars = "p[1]", #np = np, 
           n_warmup = 600, #window = c(1,900)
           ) #+ xlab("Post-warmup iteration")

```


```{r, eval=F}
#library(shinystan)
#??? ss_complete_pool <- extract(fit);

library(shinystan)
ss_output <- launch_shinystan(as.shinystan(fit, 
                              pars=c(
#              "sd_y", "mu_u1", "mu_alpha", "beta", #"rho", 
              "theta", "sd_season", "mu_season", #"sd_base", #"mu_base", 
#              "s",
              "p",
              "g", 
              "w", 
              "d", 
              "p_mu", "p_sd", "p_L", "p_err", 
              "gw_mu", "gw_sd", "gw_L", "gw_err", 
              "ab_mu", "ab_sd", "ab_L", "ab_err"
              )))

```


# "sd_y", "mu_u1", "mu_alpha", "beta", "theta", "sd_season"
```{r, fig.width=8.3, fig.height=11.7}
#```{r}

#par(mar = c(4, 4, 0.5, 0.5))
#knitr::opts_chunk$set(fig.width=6, fig.height=9)
#color_scheme_set("viridis")  
trace <- traceplot(fit, pars = c(
                                 "sd_y", "mu_u1", 
                                 "mu_alpha", "beta", 
                                 "theta", "sd_season"#,
                                 # "mu_season[1]", "mu_season[2]", "mu_season[3]"#, 
                                 # "p[1]", "p[2]", "p[3]",
                                 # "g[1]", "g[2]", 
                                 # "w[1]", "w[2]", "w[3]",
                                 # "d[1]", "d[2]", "d[3]"
                                 ), include=TRUE,
#          nrow = 2, 
          ncol = 1,
#          window=c(1,200),
#           unconstrain = FALSE,
           size = 0.5,#0.25,
           inc_warmup = TRUE
          ) + ylim(-0.2, 1.1)#3)

trace + 
  scale_color_discrete() + theme(legend.position = "top") + 
  labs(
#    x = "Weight (1000 lbs)",
    y = "parameter value",
    title = paste0("Traceplot - ", metric)
#    title = "50% posterior predictive intervals \nvs observed miles per gallon",
#    subtitle = "by vehicle weight"
  )

```

# "mu_season[1]", "mu_season[2]", "mu_season[3]", "p[1]", "p[2]", "p[3]"
```{r, fig.width=8.3, fig.height=11.7}
#```{r}

#par(mar = c(4, 4, 0.5, 0.5))
#knitr::opts_chunk$set(fig.width=6, fig.height=9)
#color_scheme_set("viridis")  
trace <- traceplot(fit, pars = c(
                                 # "sd_y", "mu_u1", 
                                 # "mu_alpha", "beta", 
                                 # "theta", "sd_season", 
                                 "mu_season[1]", "mu_season[2]", "mu_season[3]", 
                                 "p[1]", "p[2]", "p[3]"#,
                                 # "g[1]", "g[2]", 
                                 # "w[1]", "w[2]", "w[3]",
                                 # "d[1]", "d[2]", "d[3]"
                                 ), include=TRUE,
#          nrow = 2, 
          ncol = 1,
#          window=c(1,200),
#           unconstrain = FALSE,
           size = 0.5,#0.25,
           inc_warmup = TRUE
          ) + ylim(-0.2, 1.1)

trace + 
  scale_color_discrete() + theme(legend.position = "top") + 
  labs(
#    x = "Weight (1000 lbs)",
    y = "parameter value",
    title = paste0("Traceplot - ", metric)
#    title = "50% posterior predictive intervals \nvs observed miles per gallon",
#    subtitle = "by vehicle weight"
  )

```

# "g[1]", "g[2]"
```{r, fig.width=8.3, fig.height=3.9}
#```{r}

#par(mar = c(4, 4, 0.5, 0.5))
#knitr::opts_chunk$set(fig.width=6, fig.height=9)
#color_scheme_set("viridis")  
trace <- traceplot(fit, pars = c(
                                 # "sd_y", "mu_u1", 
                                 # "mu_alpha", "beta", 
                                 # "theta", "sd_season", 
                                 # "mu_season[1]", "mu_season[2]", "mu_season[3]"#, 
                                 # "p[1]", "p[2]", "p[3]",
                                 "g[1]", "g[2]"#,
                                 # "w[1]", "w[2]", "w[3]",
                                 # "d[1]", "d[2]", "d[3]"
                                 ), include=TRUE,
#          nrow = 2, 
          ncol = 1,
#          window=c(1,200),
#           unconstrain = FALSE,
           size = 0.5,#0.25,
           inc_warmup = TRUE
          ) + ylim(-0.2, 1.1)#3)

trace + 
  scale_color_discrete() + theme(legend.position = "top") + 
  labs(
#    x = "Weight (1000 lbs)",
    y = "parameter value",
    title = paste0("Traceplot - ", metric)
#    title = "50% posterior predictive intervals \nvs observed miles per gallon",
#    subtitle = "by vehicle weight"
  )

```

# "w[1]", "w[2]", "w[3]", "d[1]", "d[2]", "d[3]"
```{r, fig.width=8.3, fig.height=11.7}
#```{r}

#par(mar = c(4, 4, 0.5, 0.5))
#knitr::opts_chunk$set(fig.width=6, fig.height=9)
#color_scheme_set("viridis")  
trace <- traceplot(fit, pars = c(
                                 # "sd_y", "mu_u1", 
                                 # "mu_alpha", "beta", 
                                 # "theta", "sd_season", 
                                 # "mu_season[1]", "mu_season[2]", "mu_season[3]"#, 
                                 # "p[1]", "p[2]", "p[3]",
                                 # "g[1]", "g[2]"#,
                                 "w[1]", "w[2]", "w[3]",
                                 "d[1]", "d[2]", "d[3]"
                                 ), include=TRUE,
#          nrow = 2, 
          ncol = 1,
#          window=c(1,200),
#           unconstrain = FALSE,
           size = 0.5,#0.25,
           inc_warmup = TRUE
          ) + ylim(-0.2, 1.1)#3)

trace + 
  scale_color_discrete() + theme(legend.position = "top") + 
  labs(
#    x = "Weight (1000 lbs)",
    y = "parameter value",
    title = paste0("Traceplot - ", metric)
#    title = "50% posterior predictive intervals \nvs observed miles per gallon",
#    subtitle = "by vehicle weight"
  )

```

```{r, fig.width=11.7, fig.height=8.3}
#```{r}

color_scheme_set("brightblue") # see help("color_scheme_set")

rhat(fit) %>% 
  mcmc_rhat() + 
    labs(title = paste0("Rhat - ", metric)
         )

neff_ratio(fit) %>% 
  mcmc_neff(size = 2) + 
    labs(title = paste0("Relative N_eff - ", metric)
         )

```


# 50% posterior predictive intervals plot
```{r}
ppc_intervals(
  y = mtcars$mpg,
  yrep = posterior_predict(fit2),
  x = mtcars$wt,
  prob = 0.5
)
```


# Pairs plots
```{r} 

#par(mar = c(4, 4, 0.5, 0.5))
knitr::opts_chunk$set(fig.width=6, fig.height=6)
color_scheme_set("gray") #"darkgray")
#Error: cannot allocate vector of size 782.8 Mb
#mcmc_parcoord(posterior, np = np)

#http://mc-stan.org/bayesplot/articles/visual-mcmc-diagnostics.html
#The main problem is that large steps are required to explore the less narrow regions efficiently, but those steps become too large for navigating the narrow region. The required step size is connected to the value of tau. When tau is large it allows for large variation in theta (and requires large steps) while small tau requires small steps in theta.
#The non-centered parameterization avoids this by sampling the eta parameter which, unlike theta, is a priori independent of tau. Then theta is computed deterministically from the parameters eta, mu and tau afterwards. Here’s the same plot as above, but with eta[1] from non-centered parameterization instead of theta[1] from the centered parameterization: 


mcmc_pairs(fit.ar, np = np, pars = c("theta", "mu_season[1]", "mu_season[2]", "mu_season[3]"))

mcmc_pairs(fit.ar, np = np, pars = c("mu_u1", "sd_y", "theta", "mu_season[1]"))
mcmc_pairs(fit.ar, np = np, pars = c("mu_u1", "sd_y", "mu_season[2]", "mu_season[3]"))
mcmc_pairs(fit.ar, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "theta", "mu_season[1]"))
mcmc_pairs(fit.ar, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "mu_season[2]", "mu_season[3]"))

mcmc_pairs(fit.ar, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "theta", "mu_season[1]"))
mcmc_pairs(fit.ar, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "mu_season[2]", "mu_season[3]"))
mcmc_pairs(fit.ar, np = np, pars = c("gw_mu[3]", "gw_mu[4]", "gw_mu[5]", "theta", "mu_season[1]"))
mcmc_pairs(fit.ar, np = np, pars = c("gw_mu[3]", "gw_mu[4]", "gw_mu[5]", "mu_season[2]", "mu_season[3]"))

mcmc_pairs(fit.ar, np = np, pars = c("p_mu[1]", "p_mu[2]", "p_mu[3]", "theta", "mu_season[1]"))
mcmc_pairs(fit.ar, np = np, pars = c("p_mu[1]", "p_mu[2]", "p_mu[3]", "mu_season[2]", "mu_season[3]"))
mcmc_pairs(fit.ar, np = np, pars = c("d[1]", "d[2]", "d[3]", "theta", "mu_season[1]"))
mcmc_pairs(fit.ar, np = np, pars = c("d[1]", "d[2]", "d[3]", "mu_season[2]", "mu_season[3]"))

mcmc_pairs(fit.ar, np = np, pars = c("mu_u1", "sd_y", "ab_mu[1]", "ab_mu[2]")) 
mcmc_pairs(fit.ar, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "gw_mu[3]", "gw_mu[4]", "gw_mu[5]"))
mcmc_pairs(fit.ar, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "p_mu[1]", "p_mu[2]", "p_mu[3]"))
mcmc_pairs(fit.ar, np = np, pars = c("gw_mu[1]", "gw_mu[2]", "d[1]", "d[2]", "d[3]"))
mcmc_pairs(fit.ar, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "p_mu[1]", "p_mu[2]", "p_mu[3]"))
mcmc_pairs(fit.ar, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "d[1]", "d[2]", "d[3]"))

#====================================================
  # sd_y = 0.08, mu_u1 = y_LT = 0.1; mu_alpha = 0.04; beta = 0.6;  
# theta = 0.15; rho = 1, # sd_base = 0.2; mu_base = -1.6 
  # sd_season = 0.1, mu_season = c(-0.12, -0.06, 0.15); 
# s = c(0.4, 0.15, 0.15)  
  # p = c(0.7, 0.62, 0.68); p = c(0.7, 0.62, 0.35) [inv_logit()]; p = c(0.72, 0.42) #p[3]=0.7
  # g = c(0.85, 0.35)       #old: g = c(0.2, 0.2, 0.25) 
  # w = c(0.72, 0.15, 0.62)  
# d = c(0.05, 0.7, 0.25) 

```


# More pairs plot
```{r, eval=F}

mcmc_pairs(fit.ar, np = np, pars = c("p_sd[1]", "p_sd[2]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(fit.ar, np = np, pars = c("p_sd[1]", "p_sd[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(fit.ar, np = np, pars = c("p_sd[1]", "p_sd[2]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(fit.ar, np = np, pars = c("p_err[1]", "p_err[2]", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(fit.ar, np = np, pars = c("p_err[1]", "p_err[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(fit.ar, np = np, pars = c("p_err[1]", "p_err[2]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(fit.ar, np = np, pars = c("mu_u1", "sd_y", "p_mu[1]", "p_mu[2]", "p0[1]"))
mcmc_pairs(fit.ar, np = np, pars = c("mu_u1", "sd_y", "p_sd[1]", "p_sd[2]"))
mcmc_pairs(fit.ar, np = np, pars = c("mu_u1", "sd_y", "p_err[1]", "p_err[2]"))
 
mcmc_pairs(fit.ar, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(fit.ar, np = np, pars = c("ab_mu[1]", "ab_mu[2]", "ab_err[1]", "ab_err[2]"))
mcmc_pairs(fit.ar, np = np, pars = c("ab_sd[1]", "ab_sd[2]", "ab_err[1]", "ab_err[2]"))

mcmc_pairs(fit.ar, np = np, pars = c("mu_u1", "sd_y", "ab_mu[1]", "ab_mu[2]"))
mcmc_pairs(fit.ar, np = np, pars = c("mu_u1", "sd_y", "ab_sd[1]", "ab_sd[2]"))
mcmc_pairs(fit.ar, np = np, pars = c("mu_u1", "sd_y", "ab_err[1]", "ab_err[2]"))
mcmc_pairs(fit.ar, np = np, pars = c("mu_alpha", "beta", "mu_u1", "sd_y")) 

```


# Scatter plot
```{r, eval=F}
color_scheme_set("darkgray")
# mcmc_parcoord(fit.ar, np = np)
#
# mcmc_scatter(
#   fit.ar,
#   pars = c("p[1]", "sd_y"),
# #  transform = list(beta = "log"),
#   np = np,
#   size = 1
# )
```


```{r, eval=F}
check_hmc_diagnostics(fit)
    # check_treedepth(fit)
    # check_divergences(fit)
    # check_energy(fit)

get_max_treedepth_iterations(fit)
get_divergent_iterations(fit)
    # get_num_max_treedepth(fit)
    # get_num_divergent(fit)
    # #get_num_leapfrog_per_iteration(fit)
    # get_bfmi(fit)
    # get_low_bfmi_chains(fit)
    # get_adaptation_info(fit) # step size and full inverse mattrix

```



```{r}
sessionInfo()
```


