---
title: "try KFAS"
output: html_notebook
---

```{r}
#library(bssm)
library(KFAS) 

help(KFAS) 
```

#
y[t] = Z[t]α[t] + ε[t], (observation equation)

α[t+1] = T[t]α[t] + R[t]η[t], (transition equation)

where ε[t] ~ N(0, H[t]), η[t] ~ N(0, Q[t]) and α[1] ~ N(a[1], P[1]) independently of each other.

All system and covariance matrices Z, H, T, R and Q can be time-varying, and partially or totally missing observations y[t] are allowed.

Covariance matrices H and Q has to be positive semidefinite (although this is not checked).

Model components in KFAS are defined as

y = n x p matrix containing the observations.
Z = p x m x 1 or p x m x n array corresponding to the system matrix of observation equation.
H = p x p x 1 or p x p x n array corresponding to the covariance matrix of observational disturbances epsilon.

T = m x m x 1 or m x m x n array corresponding to the first system matrix of state equation.
R = m x k x 1 or m x k x n array corresponding to the second system matrix of state equation.
Q = k x k x 1 or k x k x n array corresponding to the covariance matrix of state disturbances eta

a1 = m x 1 matrix containing the expected values of the initial states.
P1 = m x m matrix containing the covariance matrix of the [nondiffuse] part of the initial state vector.
P1inf = m x m matrix containing the covariance matrix of the [diffuse] part of the initial state vector.

#=======================================================================================================


# Create a State Space Model Object of Class SSModel
```{r}
# add intercept to state equation by augmenting the state vector:
# diffuse initialization for the intercept, gets estimated like other states:
# for known fixed intercept, just set P1 = P1inf = 0 (default in SSMcustom).
intercept <- 0
model_int <- SSModel(Nile ~ SSMtrend(1, Q = 1469) +
SSMcustom(Z = 0, T = 1, Q = 0, a1 = intercept, P1inf = 1), H = 15099)
model_int$T
model_int$T[1, 2, 1] <- 1 # add the intercept value to level
out <- KFS(model_int)
```


```{r}
# dynamic regression model
set.seed(1)
x1 <- rnorm(100)
x2 <- rnorm(100)
b1 <- 1 + cumsum(rnorm(100, sd = 1))
b2 <- 2 + cumsum(rnorm(100, sd = 0.1))
y <- 1 + b1 * x1 + b2 * x2 + rnorm(100, sd = 0.1)
model <- SSModel(y ~ SSMregression(~ x1 + x2, Q = diag(NA,2)), H = NA)
fit <- fitSSM(model, inits = c(0,0,0), method = "BFGS")
model <- fit$model
model$Q
model$H
out <- KFS(model)
ts.plot(out$alphahat[,-1], b1, b2, col = 1:4)

```


# example of using [data] argument 
```{r}
y <- x <- rep(1, 3)
data1 <- data.frame(x = rep(2, 3))
data2 <- data.frame(x = rep(3, 3))
f <- formula(~ -1 + x)

# With data missing the environment of formula is checked,
# and if not found in there a calling environment via parent.frame is checked.
c(SSModel(y ~ -1 + x)["Z"]) # 1
c(SSModel(y ~ -1 + x, data = data1)["Z"]) # 2
c(SSModel(y ~ -1 + SSMregression(~ -1 + x))["Z"]) # 1
c(SSModel(y ~ -1 + SSMregression(~ -1 + x, data = data1))["Z"]) # 2
c(SSModel(y ~ -1 + SSMregression(~ -1 + x), data = data1)["Z"]) # 2
SSModel(y ~ -1 + x + SSMregression(~ -1 + x, data = data1))["Z"] # 1 and 2
SSModel(y ~ -1 + x + SSMregression(~ -1 + x), data = data1)["Z"] # both are 2
SSModel(y ~ -1 + x + SSMregression(~ -1 + x, data = data1), data = data2)["Z"] # 3 and 2
SSModel(y ~ -1 + x + SSMregression(f))["Z"] # 1 and 1
SSModel(y ~ -1 + x + SSMregression(f), data = data1)["Z"] # 2 and 1
SSModel(y ~ -1 + x + SSMregression(f,data = data1))["Z"] # 1 and 2
rm(x)
c(SSModel(y ~ -1 + SSMregression(f, data = data1))$Z) # 2

## Not run:
# This fails as there is no x in the environment of f
try(c(SSModel(y ~ -1 + SSMregression(f), data = data1)$Z))
## End(Not run)
```


# [<-.SSModel Extract or Replace Parts of a State Space Model
```{r}

set.seed(1)
model <- SSModel(rnorm(10) ~ 1)
model["H"]
model["H"] <- 10
# H is still an array:
model["H"]
logLik(model)
model$H <- 1
# model["H"] throws an error as H is now scalar:
model$H
logLik(model, check.model = TRUE) #with check.model = FALSE R crashes!

```





################################################################
## using deterministic inputs in observation and state equations
################################################################
```{r}
model_Nile <- SSModel(Nile ~ 
  SSMcustom(Z=1, T = 1, R = 0, a1 = 100, P1inf = 0, P1 = 0, Q = 0, state_names = "d_t") +
  SSMcustom(Z=0, T = 1, R = 0, a1 = 100, P1inf = 0, P1 = 0, Q = 0, state_names = "c_t") +
  SSMtrend(1, Q = 1500), H = 15000)
model_Nile$T
model_Nile$T[1, 3, 1] <- 1 # add c_t to level
model_Nile0 <- SSModel(Nile ~ 
  SSMtrend(2, Q = list(1500, 0), a1 = c(0, 100), P1inf = diag(c(1, 0))), 
  H = 15000)

ts.plot(KFS(model_Nile0)$mu, KFS(model_Nile)$mu, col = 1:2)
```


##########################################################
### Examples of generalized linear modelling with KFAS ###
##########################################################
```{r}
# Same example as in ?glm
counts <- c(18, 17, 15, 20, 10, 20, 25, 13, 12)
outcome <- gl(3, 1, 9)
treatment <- gl(3, 3)
glm_D93 <- glm(counts ~ outcome + treatment, family = poisson())

model_D93 <- SSModel(counts ~ outcome + treatment,
  distribution = "poisson")

out_D93 <- KFS(model_D93)
coef(out_D93, last = TRUE)
coef(glm_D93)

summary(glm_D93)$cov.s
out_D93$V[, , 1]

# approximating model as in GLM
out_D93_nosim <- KFS(model_D93, smoothing = c("state", "signal", "mean"))

# with importance sampling. Number of simulations is too small here,
# with large enough nsim the importance sampling actually gives
# very similar results as the approximating model in this case
set.seed(1)
out_D93_sim <- KFS(model_D93,
  smoothing = c("state", "signal", "mean"), nsim = 1000)


## linear predictor
# GLM
glm_D93$linear.predictor
# approximate model, this is the posterior mode of p(theta|y)
c(out_D93_nosim$thetahat)
# importance sampling on theta, gives E(theta|y)
c(out_D93_sim$thetahat)


## predictions on response scale
# GLM
fitted(glm_D93)
# approximate model with backtransform, equals GLM
fitted(out_D93_nosim)
# importance sampling on exp(theta)
fitted(out_D93_sim)

# prediction variances on link scale
# GLM
as.numeric(predict(glm_D93, type = "link", se.fit = TRUE)$se.fit^2)
# approx, equals to GLM results
c(out_D93_nosim$V_theta)
# importance sampling on theta
c(out_D93_sim$V_theta)


# prediction variances on response scale
# GLM
as.numeric(predict(glm_D93, type = "response", se.fit = TRUE)$se.fit^2)
# approx, equals to GLM results
c(out_D93_nosim$V_mu)
# importance sampling on theta
c(out_D93_sim$V_mu)
```



## Not run: 
####################################
### Linear mixed model with KFAS ###
####################################
```{r}
# example from ?lmer of lme4 pacakge
data("sleepstudy", package = "lme4")

model_lmm <- SSModel(Reaction ~ Days +
    SSMregression(~ Days, Q = array(0, c(2, 2, 180)),
       P1 = matrix(NA, 2, 2), remove.intercept = FALSE), sleepstudy, H = NA)

# The first 10 time points the third and fouth state
# defined with SSMregression correspond to the first subject, and next 10 time points
# are related to second subject and so on.

# need to use ordinary $ assignment as [ assignment operator for SSModel
# object guards against dimension altering
model_lmm$T <- array(model_lmm["T"], c(4, 4, 180))
attr(model_lmm, "tv")[3] <- 1L #needs to be integer type!

# "cut the connection" between the subjects
times <- seq(10, 180, by = 10)
model_lmm["T",states = 3:4, times = times] <- 0

# for the first subject the variance of the random effect is defined via P1
# for others, we use Q
model_lmm["Q", times = times] <- NA

    # data("GlobalTemp")
    # inits <- chol(cov(GlobalTemp))[c(1, 4, 3)]
    # inits[1:2] <- log(inits[1:2])

update_lmm <- function(pars = init, model){
  P1 <- diag(exp(pars[1:2]))
  P1[1, 2] <- pars[3]
  model["P1", states = 3:4] <- model["Q", times = times] <-
    crossprod(P1)
  model["H"] <- exp(pars[4])
  model
} 

inits <- c(0, 0, 0, 3)

fit_lmm <- fitSSM(model_lmm, inits, update_lmm, method = "BFGS")
out_lmm <- KFS(fit_lmm$model)
# unconditional covariance matrix of random effects
fit_lmm$model["P1", states = 3:4]

# conditional covariance matrix of random effects
# same for each subject and time point due to model structure
# these differ from the ones obtained from lmer as these are not conditioned
# on the fixed effects
out_lmm$V[3:4,3:4,1]

```
## End(Not run)







#=======================================================================================================
#=======================================================================================================





```{r}


```

