---
title: "State space model of accounting manipulation - Stan"
output: html_notebook
---

```{r}

#library(tidyverse)
#library(dplyr)
library(magrittr)

#load previously saved Stan model objects
#load(file = "StanObjects.rda")

library(rstan)
rstan_options(auto_write = TRUE) # avoid recompilation of unchanged Stan programs

# The following can affect distributional computing over a cluster
options(mc.cores = parallel::detectCores())  # 

# The following throws an error in compiling .stan
#Sys.setenv(LOCAL_CPPFLAGS = '-march=native') # For improved execution time but can throw errors for some CPUs

```


# Estimate the simulated model with Stan - misreporting extent 
  (forecasts specified in generated quantities block)
//ignore these lines: 
//  COGS2Rev, SGA2Rev, RnD2Rev with Rev2Rev = unit vector as the intercept)
//   int<lower=0,upper=1> M[N];   // M = 1 for the decision to misreport; = 0 if report honestly  

```{stan output.var=SSM, eval=T}
// Focus on the time series of only one firm
// The observations are reported accounting figures, say Gross Profit, (scaled by Rev, total revenue) a vector r of length N (years).
// The covariates are a matrix capital X with K column vectors, incl a unit vector, capturing the  
//   dis/incentives of misreporting the underlying unbiased figures (y) as the observed reported firgures
data {
  int<lower=0> N; // number of observations (for different years)
  int<lower=0> K; // number of coefficents for the column vectors in covariate matrix X
  vector[N] r; // reported figures (scaled by Rev, total revenue)
  matrix[N,K] X; // covariates (capturing the dis/incentives of misreporting)

// forecasts  
//  int<lower=0> N_new; // number of predictions
//  matrix[N_new,K] X_new; // 
}
transformed data {
//  real <lower=0,upper=1> a = 0; // intercept coefficient (drift) of the AR(1) process of the unbiased figure y[n]
// GP data
}
parameters {
  vector[N] y; // underlying unbiased figure (scaled by Rev, total revenue)
  real <lower=0,upper=1> a; // intercept coefficient (drift) of the AR(1) process of the unbiased figure y[n]
  real <lower=0,upper=1> b; // slope coefficient of the AR(1) process of y[n]
  vector[K] c; // coefficients of the K covariates in matrix X
  real<lower=0> sd_y; // sd of the underlying unbiased figures (vector y)
  real<lower=0> sd_m; // sd of the misreporting extents (vector m = r - y)
// real<lower=0,upper=1> integrity;     // integrity level of the society affecting the decision to misreport or not
//  vector[N] integrity;     // integrity level of each CEO affecting the decision to misreport or not
}
transformed parameters {
//  vector[N] m; // misreporting extent in the reported figures 
  vector[N] m = r - y;

//  real ilogodds;     // log odds of the society integrity level
//  ilogodds = logit(integrity);   
}
model {

  y[1] ~ normal(a/(1-b), sd_y);   // y should be nonnegative for Rev and nonpositive for Costs
//  for (n in 2:N)
//    y[n] ~ normal(alpha + beta * y[n-1], sigma);
  y[2:N] ~ normal(a + b * y[1:(N - 1)], sd_y);
  
// c ~ cauchy(0, 2.5); // common prior for each b[K]
  m ~ normal(X*c, sd_m);  // shouldn't good governance restrict |m| or m^2 ? Then impact of X is on sd_m
//  M ~ bernoulli_logit(ilogodds); 

}
generated quantities {

//  vector[N_new] r_new;
//  for (n in 1:N_new)
//    r_new[n] = normal_rng(Z_new[n] * b, sd_r);
}

```

# Simulate data
```{stan output.var=simu_SSM, eval=T}
data {
  int<lower=0> N; // number of observations (for different years)
  int<lower=0> K; // number of coefficents for the column vectors in covariate matrix X
  matrix[N,K] X; // covariates (capturing the dis/incentives of misreporting)

  real a; // intercept coefficient (drift) of the AR(1) process of the underlying unbiased accounting figure y[n]
  real b; // slope coefficient of the AR(1) process of y[n]
  vector[K] c; // coefficients of the K covariates in matrix X
  real<lower=0> sd_y; // sd of the underlying unbiased figure (vector y)
  real<lower=0> sd_m; // sd of the misreporting extent (vector m = r - y)
}
transformed data {
//  real a = 0;
}
parameters {}
model {}
generated quantities {
  vector[N] y; // underlying unbiased figures (eg, Gross Profit, scaled by Rev)
  vector[N] m; // misreporting extent in the reported figures 
//  vector[N] r; // reported figures (scaled by Rev, total revenue)

  y[1] = normal_rng(a/(1-b), sd_y);   // y should be nonnegative for Rev and nonpositive for Costs
// Perhaps best to model y as always positive, with IsCost = 1, 0 to indicate Cost or Rev item 
//  y[2:N] = normal_rng(a + b * y[1:(N - 1)], sd_y);
//  m = normal_rng(X*c, sd_m);
//  r = y + m; 
  m[1] = normal_rng(X[1]*c, sd_m); 
//  r[1] = y[1] + m[1];     // should be more efficient to use R to create r from y, m
  for (n in 2:N) {
    y[n] = a + normal_rng(b * y[n-1], sd_y);
    m[n] = normal_rng(X[n]*c, sd_m); 
//    r[n] = y[n] + m[n];
  }
}
```


# Simulate data and fit the debug model (also, save Stan model objects) 
```{r}

N = 16*5  #7 number of (yearly) observations of a firm

sd_y = 0.02  # sd of underlying unbiased figures is likely to be industry-specific / business model specific
sd_m = 0.01  # sd of misreporting extent m, which should be restrained by the governance setting (eg, audit committee chair with financial expertise, big 4 auditor, etc)

a = 0#.04  # intercept parameter of the AR(1) processs of the underlying unbiased figure
b = 0.6  # slope parameter of the AR(1) processs of the underlying unbiased figure
#set parameters of the data generation process of the misreporting extent m
c = c(-0.07, 0.2)
#c = c(0.1, -0.07, 0.2)
K = length(c)
cvec <- matrix(data = c, nrow = K)  # define column vector as a matrix[N,1]
d = 1 # fractional reversal of prior-period manipluation by accruals

integrity = 0.8       
# integrity level, between (0,1), of the society, interpreted as likelihood to behave opportunistically
# integrity level of an individual, indiv_integ between (-Inf,Inf), is 
#   affected by a society's general integrity level in a stochastic fashion varying from period to period

#M <- rbinom(N, 1, integrity)   # the decision to misreport earnings

#set.seed(88)  # set the seed for rng to ensure replicability
# covariate matrix X capturing the dis/incentive to misreport
X <- matrix(data = c(#rep(1, N), 
                     rnorm(N, integrity, 0.3), rbinom(N, 1, 0.3)),
            nrow = N)
colnames(X) <- c(#"constant", 
                 "indiv_integ", "downgrade")  
# downgrade = 1 if one or more analysts following the firm downgraded their stock recommendation in the last period
#colnames(Z) <- c("Rev2Rev", "COGS2Rev", "SGA2Rev", "RnD2Rev")

simu_SSM_data <- list(N = N, 
                      K = K,
                      X = X,
                      a = a,
                      b = b, 
                      c = c,
                      sd_y = sd_y,
                      sd_m = sd_m
                      )
save(simu_SSM_data, file="simu_SSM_data.rda")
load(file="simu_SSM_data.rda")
# Simulate SSM data
simu_fit <- stan(#model_code = simu_SSM@model_code,
                 file = "simu_SSM.stan",  # Stan program
                 data=simu_SSM_data, iter=1, #seed=987,  
                 chains=1, algorithm="Fixed_param")  

y <- extract(simu_fit)$y[1,]
m <- extract(simu_fit)$m[1,]
D <- m - d * dplyr::lag(m)
D[1] <- m[1]
r <- y + D

# Prepare SSM data for estimation using MCMC
#N_new = 3
SSM_data <- list(N = N, 
                 K = K,
                 r = r, 
                 X = X
                 # Z_new = Z[1:N_new,],
                 # N_new = N_new
                 )
# Fit the debug model to make sure the Stan code complies properly 
#fit0_debug <- stan(model_code = SSM@model_code, data = SSM_data, iter = 10, chains = 1)

#save(simu_GP, misreport, file = "StanObjects.rda")
#load(file = "StanObjects.rda")
```

# Fit the full model (by referring to the debug model) to estimate underlying parameters
```{r}

# Run the full model and refer to the debug model to save compilation time 
system.time({
fit1 <- stan( 
  file = "SSM.stan",  # Stan program
  data = SSM_data,    # named list of data
#  model_code = SSM@model_code, 
#  fit = fit0_debug,   # to save compilation time if the debug model was run
  control = list(adapt_delta = 0.999, max_treedepth = 12),    # adjust when there're divergent transitions after warmup
#  chains = 1,             # default is 4 Markov chains
#  cores = 8,
#  init = init, # where init = list(list(mu=...,sigma=...), list(mu=..., sigma=...), ...) 
                # where length of list = number of chains
#  seed = 555,
  iter = 9000, #500 #2000,            # total number of iterations per chain
  warmup = 7000,          # number of warmup iterations per chain
  refresh = 1000#0          # = 0 means no progress shown
  )
})


#extract(fit1)$y[1,]
#pairs(extract(fit1), pars = c(a, b)) #, y[1], sd_y, m[1], sd_m, c[1], c[2], c[3]))

```

# Summary of the posterior sample 
```{r}

ss_complete_pool <- extract(fit1);
# rho_T <- 5.5, alpha_T <- 3, sigma_T <- 2
print(fit1, probs = c(0.1, 0.5, 0.9), digits = 3) #, probs = c(.05,.95))

source('stan_utility.R')
check_all_diagnostics(fit1)

```

```{r}
warnings()
sessionInfo()
```



